
---
# This RMD file loads the cellranger count data and performs single-cell gene expression analysis
title: "Single-Cell naive and eye cells monocle2"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: no
---
  
```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
options(knitr.duplicate.label = "allow")
knitr::opts_chunk$set()
```

## 1. Introduction
This document takes the cellranger count data for each of the samples, does qc anlaysis, filtering, generates before and after qc plots and combines the data.
```{r packages, message=FALSE, warning=FALSE}
#BiocManager::install(c("Biobase","BiocGenerics","DelayedArray","DelayedMatrixStats","limma","S4Vectors","SingleCellExperiment","SummarizedExperiment"))
devtools::install_version('speedglm', '0.3-4', repos='https://packagemanager.rstudio.com/cran/2023-03-31')
install.packages("devtools")
#devtools::install_github('cole-trapnell-lab/monocle3')

install.packages("hdf5r")
library(Seurat)
library(ggplot2)
library(cowplot)
library(tidyverse)
library(viridis)
library(tradeSeq)
library(RColorBrewer)
library(HGNChelper)
library(EnhancedVolcano)
library(ComplexHeatmap) 
#library(monocle)
library(patchwork)
```

## 2. Data setup
```{r datasetup, message=FALSE, warning=FALSE}
# Set the input, output paths
directorypath="/Users/pengz4/Desktop/2023.1 10x data analysis/2023.1.4 Vijay Results/ZixuanNagarajan/Results/cellranger/"
outputpath="/Users/pengz4/Desktop/2023.1 10x data analysis/2023.2.10 EyeCell and Naive output/"
sampleinfopath="/Users/pengz4/Desktop/2023.1 10x data analysis/2023.1.4 Vijay Results/ZixuanNagarajan/SampleInformation.tab"
samples=c("S1_HTO_1","S1_HTO_2","S1_HTO_3","S1_HTO_4","S3_HTO_1")
setwd("/Users/pengz4/Desktop/2023.1 10x data analysis/2023.2.10 EyeCell and Naive output/")
```
## 3. Quality Analysis and Filtering
This part of the workflow reads in each of the samples, analyses the quality, performs the quality filter and merges the data. Both merged unfiltered and merged filtered data are then used for plotting the quality
```{r qc, message=FALSE, warning=FALSE}
# Using the for loop to reads the samples list
# Create empty lists to hold the combined data set
allsamplestomerge=list()
allsamplestomergeprefilter=list()

# Using the for loop to reads the samples list
for(i in samples)
{
  # Remove any old variables
  rm(onesampledataFilt)
  rm(onesampledata)
  rm(onesample)
  # Print the current sample name
  print(i)
  # Read the data file
  onesample <- Read10X_h5(paste0(directorypath,paste0(i,"/sample_filtered_feature_bc_matrix.h5",sep=""),sep=""), use.names = T)

  # Print column names
  colnames(onesample$`Gene Expression`)
  # Add gene names to columnnames
  colnames(onesample$`Gene Expression`) <-     paste(sapply(strsplit(colnames(onesample$`Gene Expression`),split="-"),'[[',1L),i,sep="-")
  # Print column names
  colnames(onesample$`Gene Expression`)
  
  # Create a Seurat data object from the gex matrix, keeping all the features, and only cells with at least 300 genes
  onesampledata <- CreateSeuratObject(counts = onesample[["Gene Expression"]], min.cells = 0,min.features = 300,names.field = 2,names.delim = "\\-")
  
  # Check genes names
  head(rownames(onesampledata))
  # Check data rows and columns
  head(onesampledata)
  
  # Generate Percent mito for each of the cells
  # Lookup genes
  grep("^mt-",rownames(onesampledata@assays$RNA@counts),value = TRUE)
  grep("Rp[sl]",rownames(onesampledata@assays$RNA@counts),value = TRUE)
  grep("^Hb[^(p)]",rownames(onesampledata@assays$RNA@counts),value = TRUE)
  
  # Generate Percent mito for each of the cells
  onesampledata$percent.mito <- PercentageFeatureSet(onesampledata, pattern = "^mt-")
  summary(onesampledata$percent.mito)
  head(onesampledata)
  
  # Generate Percent ribo for each of the cells
  onesampledata$percent.ribo <- PercentageFeatureSet(onesampledata, pattern = "Rp[sl]")
  summary(onesampledata$percent.ribo)
  head(onesampledata)
  
  # Generate Percent hemoglobin for each of the cells
  onesampledata$percent.hb <- PercentageFeatureSet(onesampledata, pattern = "^Hb[^(p)]")
  summary(onesampledata$percent.hb)
  head(onesampledata)
  
  # Check the total number of cells in this sample
  table(onesampledata$orig.ident)
  # Filter data and keep only cells that has percent.mito less than or equal to 8
  onesampledataFilt <- subset(onesampledata, percent.mito <= 8)
  # Check the number of cells after mito filtering
  table(onesampledataFilt$orig.ident)
  # Filter for nCount RNA (total rna molecules sequenced)
  onesampledataFilt <- subset(onesampledataFilt, nCount_RNA >= 1000 & nCount_RNA <= 40000)
  # Check number of cells after ncount filtering
  table(onesampledataFilt$orig.ident)
  # Filter for nfeature (total number of genes expressed)
  onesampledataFilt <- subset(onesampledataFilt, nFeature_RNA >= 400)
  # Check number of cells after nfeature filtering
  table(onesampledataFilt$orig.ident)
  TRUE# Filter for percent ribo
  onesampledataFilt <- subset(onesampledataFilt, percent.ribo > 10 & percent.ribo < 45)
  # Counts after filtering
  table(onesampledataFilt$orig.ident)
  
  # Add filtered and unfiltered data to a list
  allsamplestomerge[[i]]=onesampledataFilt
  allsamplestomergeprefilter[[i]]=onesampledata
}
```
## 4. Merge samples
```{r merge, message=FALSE, warning=FALSE}
# Create one and the rest of the samples as a list, for filtered data
one=allsamplestomerge[[1]]
one
rest=allsamplestomerge[2:length(allsamplestomerge)]
rest

# Create one and the rest of the samples as a list, for profiltered data
onepre=allsamplestomergeprefilter[[1]]
onepre
restpre=allsamplestomergeprefilter[2:length(allsamplestomergeprefilter)]
restpre

# Merge filtered seurat objects
allsamples <- merge(one, y = rest)
allsamples
head(allsamples)
table(Idents(allsamples))

# Merge prefiltered seurat objects
allsamplespre <- merge(onepre, y = restpre)
allsamplespre
head(allsamplespre)
table(Idents(allsamplespre))
```
## 5. Add metadata
```{r metadata, message=FALSE, warning=FALSE}
## Assign sample types metadata column
 allsamples$groups <- plyr::mapvalues(
  x = allsamples$orig.ident, 
  from = c("S1_HTO_1","S1_HTO_2","S1_HTO_3","S1_HTO_4","S3_HTO_1"),
  to = c("S1","S1","S1","S1","S3"))
head(allsamples)

allsamples$type <- plyr::mapvalues(
  x = allsamples$orig.ident, 
  from = c("S1_HTO_1","S1_HTO_2","S1_HTO_3","S1_HTO_4","S3_HTO_1"),
  to = c("eye","eye","eye","eye","Naive T"))
head(allsamplespre)

# View type counts
table(allsamples$type)
table(allsamples$groups)
# View data
head(allsamples)
```
## 6. Generate before and after QC plots
```{r plots, message=FALSE, warning=FALSE}
# Generate before filtering quality plots
qcbefore=VlnPlot(allsamplespre,features = c("nFeature_RNA", "nCount_RNA","percent.mito","percent.ribo"),ncol = 2, pt.size = 0) +
  NoLegend()
qcbefore
# Save the plot in current working directory
ggsave(paste(outputpath,"before-qc-violineplot.pdf",sep=""), plot = qcbefore, width = 15, height = 20, units = "cm")

# Generate after filtering quality plots
Idents(allsamples) <- "orig"
qcafter=VlnPlot(allsamples,features = c("nFeature_RNA", "nCount_RNA","percent.mito","percent.ribo"),ncol = 2, pt.size = 0) +
  NoLegend()
qcafter
Idents(allsamples) <- "orig.ident"
head(allsamples)
# Save the plot in current working directory
ggsave(paste(outputpath,"after-qc-violineplot.pdf",sep=""), plot = qcafter, width = 15, height = 20, units = "cm")
```
## 7. Normalization
```{r normalize, message=FALSE, warning=FALSE}
###################### Normalize Data
# Feature counts for each cell are divided by the total counts for that cell and multiplied by the scale.factor. This is then natural-log transformed
# Cell level normalization - accounts for sequencing depth
allsamples <- NormalizeData(
  object = allsamples,
  normalization.method = "LogNormalize",
  scale.factor = 10000)

##################### Find variable genes
# Returns top 2000 variable genes
# VST is uses LOESS method
# FindVariableFeatures needs normalization
allsamples <- FindVariableFeatures(
  object = allsamples,
  selection.method = "vst")
top10 <- head(VariableFeatures(allsamples), 10)
top10
# Plot variable features and label the top 10
vfp1 <- VariableFeaturePlot(allsamples)
vfp1 <- LabelPoints(plot = vfp1, points = top10, repel = TRUE)
vfp1
ggsave(paste(outputpath,"variablefeatureplot.png",sep=""), plot = vfp1, width = 20, height = 15, units = "cm")
```
## 8. Scaling & Batch Correction
```{r scale, message=FALSE, warning=FALSE}
##################### SCALE DATA
# PCA needs scaled data
# Zero centers and scales (mean/sd) gene/feature data in each cell (for across sample comparison), so extreme ranges in expression do not affect clustering (done for making cells with similar expression cluster together)
dim(allsamples)
allsamples <- ScaleData(
  object = allsamples,
  # Remove unwanted sources of variation (technical, batch etc.)
  vars.to.regress = c("percent.mito", "nFeature_RNA", "percent.ribo", "nCount_RNA"))
```
## 9. PCA
```{r PCA, message=FALSE, warning=FALSE}
##################### PCA
# Run PCA with 50 pcs
allsamples <- RunPCA(object = allsamples, npcs=50)
# Print genes in the top 5 pcs
print(allsamples[["pca"]], dims = 1:5, nfeatures = 5)
# Identify number of PCs that explains majority of variations
ElbowPlot(allsamples, ndims=50, reduction = "pca")
# Visualize the genes in the first PC
VizDimLoadings(allsamples, dims = 1, ncol = 1) + theme_minimal(base_size = 8)
# Visualize the genes in the second PC
VizDimLoadings(allsamples, dims = 2, ncol = 1) + theme_minimal(base_size = 8)
# Set the identity column to WT vs TG
Idents(allsamples) <- "orig.ident"
# Plot heatmap with cells=500 plotting cells with extreme cells on both ends of spectrum
#DimHeatmap(object = allsamples, dims = 1:6, cells = 500, balanced = TRUE)
DimHeatmap(object = allsamples, dims = 1:6, cells = 500, balanced = TRUE)
```
## 10. Clustering
```{r clustering, message=FALSE, warning=FALSE}
##################### Clustering
# use the first 20 pc's
use.pcs = 1:20
# FindNeighbors with 20 pcs
allsamples <- FindNeighbors(allsamples, reduction="pca", dims = use.pcs)
# FindClusters with resolution starting at 0.20 and ending at 2, with 0.1 increment
allsamples <- FindClusters(
  object = allsamples,
  resolution = seq(0.20,0.65,0.15),
  verbose = FALSE)
head(allsamples)
# Count number of clusters at each resolution
sapply(grep("res",colnames(allsamples@meta.data),value = TRUE),
       function(x) length(unique(allsamples@meta.data[,x])))
```
## 11. UMAP 
```{r umap, message=FALSE, warning=FALSE}
########################## UMAP
# Run umap with 20 pcs
set.seed(20)
allsamples <- RunUMAP(
  object = allsamples,
  reduction.use = "pca",
  dims = use.pcs)
# Plot umap
Idents(allsamples) <- "RNA_snn_res.0.2"
DimPlot(object = allsamples, pt.size=0.2, reduction = "umap", label = T)

# list cell number in each cluster for sample types
table(Idents(allsamplesclean3),allsamplesclean$type)

head(allsamplesclean3)
table(Idents(allsamplesclean3),allsamplesclean$orig.ident)
table(Idents(eyenaive),eyenaive$orig.ident)

propotioninfo=table(Idents(eyenaive),eyenaive$orig.ident)
write.table(propotioninfo, file=paste(outputpath,"propotioninfo4.txt",sep=""), quote=F, sep = "\t")
# Save the file of cell number in each cluster text file
write.table(propotioninfo, file=paste(outputpath,"propotioninfor0.5.txt",sep=""), quote=F, sep = "\t")
```
## 12 Subset:allsamplesclean and re-PCA, cluster
```{r differentialbetweeneveryclusters, message=FALSE, warning=FALSE}
###########################
# Remove "Macrophages" in r=0.2
Idents(allsamples) <- "RNA_snn_res.0.2"
table(Idents(allsamples))
allsamplesclean2 <- subset(allsamples, idents = 5 , invert = TRUE)
table(Idents(allsamplesclean2))
DimPlot(allsamplesclean2, reduction = "umap", label = T, repel = TRUE) 

# clean3, re-PCA.
allsamplesclean3 <- subset(allsamples, idents = 5 , invert = TRUE)
# Run PCA with 50 pcs
allsamplesclean3 <- RunPCA(object = allsamplesclean3, npcs=50)
# Print genes in the top 5 pcs
print(allsamplesclean3[["pca"]], dims = 1:5, nfeatures = 5)
# Identify number of PCs that explains majority of variations
ElbowPlot(allsamplesclean3, ndims=50, reduction = "pca")
# Visualize the genes in the first PC
VizDimLoadings(allsamplesclean3, dims = 1, ncol = 1) + theme_minimal(base_size = 8)
# Visualize the genes in the second PC
VizDimLoadings(allsamplesclean3, dims = 2, ncol = 1) + theme_minimal(base_size = 8)

### Re-cluster
allsamplesclean3 <- FindNeighbors(allsamplesclean3, reduction="pca", dims = use.pcs)
allsamplesclean3 <- FindClusters(
  object = allsamplesclean3,
  resolution = seq(0.20,0.65,0.025),
  verbose = FALSE)
head(allsamplesclean3)
# Count number of clusters at each resolution
sapply(grep("res",colnames(allsamplesclean3@meta.data),value = TRUE),
       function(x) length(unique(allsamplesclean3@meta.data[,x])))

# Run umap with 20 pcs
set.seed(20)
allsamplesclean3 <- RunUMAP(
  object = allsamplesclean3,
  reduction.use = "pca",
  dims = use.pcs)
Idents(allsamplesclean3) <- "RNA_snn_res.0.2"
DimPlot(allsamplesclean3, reduction = "umap", label = TRUE, repel = TRUE)
DimPlot(object = allsamplesclean3, pt.size=0.2, reduction = "umap", group.by = "type",label = T)
DimPlot(object = allsamplesclean3, pt.size=0.2, reduction = "umap", split.by = "orig.ident")

table(Idents(allsamplesclean3))

DimPlot(object = eyenaive, pt.size=0.2, reduction = "umap", group.by = "type",cols = c("#2dc2a9","salmon"))
Idents(eyenaive) <- "eyenaive_annotation"
DimPlot(object = eyenaive, pt.size=0.2, reduction = "umap")
## Assign sample types metadata column 2023.5.14
eyenaive$eyenaive_annotation=recode(as.character(eyenaive$RNA_snn_res.0.2),
                              "0"="anergic_naive",
                              "1"="anergic_IL17",
                              "2"="naiveT",
                              "3"="Treg",
                              "4"="proliferative")

head(eyenaive)
Idents(eyenaive) <- "eyenaive_annotation"
levels(eyenaive)
DimPlot(object = eyenaive, pt.size=0.2, reduction = "umap",label = T)

length(WhichCells(eyenaive, expression = Foxp3 > 0 , idents = "Treg"))
#1245
length(WhichCells(eyenaive, expression = Foxp3 > 0 , idents = "anergic_IL17"))
#430
```

## 13## Subset eyecells (2023.4.17 RE-RUN Only subset)
```{r reclustering, message=FALSE, warning=FALSE}
##subset eye cells and rePCA Umap
eyecells <- subset(allsamplesclean2, idents = c(0,2,3,4))
DimPlot(eyecells, reduction = "umap", label = TRUE, repel = TRUE) 

# Run umap with 20 pcs
set.seed(20)
eyecells <- RunUMAP(
  object = eyecells,
  reduction.use = "pca",
  dims = use.pcs)
# Plot umap
Idents(eyecells) <- "RNA_snn_res.0.2"
DimPlot(object = eyecells, pt.size=0.2, reduction = "umap", label = T)
table(Idents(eyecells))
DimPlot(object = eyecells, pt.size=0.2, reduction = "umap", label = T, cols=c("salmon", "#22c981","#23acde", "#de6de8"))


DimPlot(allsubset_eye, reduction = "umap", label = TRUE, repel = TRUE) 
DimPlot(allsamplesclean3, reduction = "umap", label = TRUE, repel = TRUE)
Idents(eyecells) <- "RNA_snn_res.0.275"
table(Idents(eyecells))
head(allsamplesclean3)

## Assign sample types metadata column 2023.5.14注释 code from Ming
eyecells$eyecells_annotation=recode(as.character(eyecells$RNA_snn_res.0.275),
                              "0"="anergic_naive",
                              "2"="Treg",
                              "3"="anergic_IL17",
                              "4"="proliferative",
                              "5"="anergic_IFNg")

# allsubset_eye$'Bach1(+)'=NULL. ###Delete any column 2023.5.15
head(eyecells)
```

## 13 FindAllMarkers (subset:allsamplesclean3)
```{r differentialbetweeneveryclusters, message=FALSE, warning=FALSE}
###########################
markersall3=FindAllMarkers(allsamplesclean3)
head(markersall3)
write.table(markersall3, file=paste(outputpath,"markersall3.txt",sep=""), quote=F, row.names = FALSE, sep = "\t")
markersall3filtered=dplyr::filter(markersall3, p_val_adj < 0.05 & (avg_log2FC > 1 | avg_log2FC < -1))
nrow(markersall3filtered)
nrow(markersall3)
write.table(markersall3filtered, file=paste(outputpath,"markersall3filtered.txt",sep=""), quote=F, row.names = FALSE, sep = "\t")

# Extract the top 10 marker for each cluster and plot heatmap (2023.4.14)
markersall3 %>%
  group_by(cluster) %>%
  top_n(n = 10, wt = avg_log2FC) -> top10
top10
# Get only the marker - gene names
all_markers<- top10 %>%
  select(gene) %>% 
  unclass() %>% 
  stack() %>%
  pull(values) %>%
  unique() %>%
  .[!is.na(.)]
all_markers

# Scale all genes - not just variable genes (2023.5.24)
allsamplesclean3heatmap = allsamplesclean3
allsamplesclean3heatmap <- ScaleData(
  object = allsamplesclean3,
features = rownames(allsamplesclean3),
# Remove unwanted sources of variation (technical, batch etc.)  
vars.to.regress = c("percent.mito", "nFeature_RNA", "percent.ribo", "nCount_RNA"))

# heatmap Fig.S1
DoHeatmap(allsamplesclean3heatmap, features = all_markers, group.by = "clean2_annotation") + NoLegend()
## make the black color map to 0. the yellow map to highest and the purle map to the lowest
DoHeatmap(allsamplesclean3heatmap, features = all_markers, group.by = "RNA_snn_res.0.2") + scale_fill_gradientn(colors = c("steelblue4", "white", "firebrick3"))+  NoLegend()
```

## 14. FeaturePlot to check markers 
```{r FeaturePlot, message=FALSE, warning=FALSE}
########################### tables
# reorder the cluster number.2023.7.16
Idents(allsamplesclean3) <- "RNA_snn_res.0.2"
levels(allsamplesclean3)
DimPlot(object = allsamplesclean3, pt.size=0.2, reduction = "umap")
eyenaive<-allsamplesclean3
Idents(eyenaive) <- factor(Idents(allsamplesclean3),levels=c(2,0,1,3,4))

Idents(eyenaive) <- "RNA_snn_res.0.2"
DimPlot(object = eyenaive, pt.size=0.2, reduction = "umap")

FeaturePlot(eyenaive, features=c("Cd44"),)
FeaturePlot(eyenaive, features=c("Mki67"),)
FeaturePlot(eyenaive, features=c("Sell"),)
FeaturePlot(eyenaive, features=c("Foxp3"),)

FeaturePlot(eyenaive, features=c("Tbx21"),)
FeaturePlot(eyenaive, features=c("Rorc"),)
FeaturePlot(eyenaive, features=c("Gata3"),)#Th2
FeaturePlot(eyenaive, features=c("Cxcr5"),)

FeaturePlot(eyenaive, features=c("Itga2"),) #CD49b Tr1
FeaturePlot(eyenaive, features=c("Irf4"),)
#Tr1 Reference paper: https://doi.org/10.1016/j.immuni.2018.12.001 Immunity
FeaturePlot(eyenaive, features=c("Ccr8"),)
FeaturePlot(eyenaive, features=c("Ccl7"),) #Th22
FeaturePlot(eyenaive, features=c("Ccl15"),)
FeaturePlot(eyenaive, features=c("Il2"),)
FeaturePlot(eyenaive, features=c("Ifng"),)
FeaturePlot(eyenaive, features=c("Il17a"),)
FeaturePlot(eyenaive, features=c("Csf2"),)

FeaturePlot(eyenaive, features=c("Il10"),)


FeaturePlot(eyenaive, features=c("Il22"),)
FeaturePlot(eyenaive, features=c("Tgfb2"),)#only Tgfb1 out of 123
FeaturePlot(eyenaive, features=c("Il12a"),)#Il35
FeaturePlot(eyenaive, features=c("Ebi3"),)#Il35

FeaturePlot(eyenaive, features=c("Foxp3","Il10"),pt.size=0.2,cols=c ("#2f7bf5","#f25b24"),blend=T,blend.threshold = 0.1)

# naive related marker: Tcf7, Ccr7, Aff3, Il7r, #Lef1 #Foxo1

FeaturePlot(eyenaive, features=c("Havcr2"),)#Tim-3
FeaturePlot(eyenaive, features=c("Izumo1r")) #FR4
FeaturePlot(eyenaive, features=c("Nt5e")) #CD73
FeaturePlot(eyenaive, features=c("Entpd1"),)#CD39 
FeaturePlot(eyenaive, features=c("Pecam1"),)#CD31
FeaturePlot(eyenaive, ffeatures=c("Itga2"),)#CD49b no expression
FeaturePlot(allsamplesclean3, features=c("Il1r1"),)
# 'Rara','Rarb','','Rarg',

#ATF6,MALT1,NFKB1,NFKBIA,PIK3CD,PRKCB,PIK3CD,PIK3R5,PRKAR1A Dgka,Dgkz
#PKCθ Signaling: Card11,CD80,GRAP2,MALT1,NFKB1,NFKBIA,Pik3cd,Pik3r5,Prkcq

FeaturePlot(eyenaive, features=c("Nfatc1"),) #NFAT 
FeaturePlot(eyenaive, features=c("Irf4"),) #IRF4 
FeaturePlot(eyenaive, features=c("Ndrg1"),) #CD45
featurecolor <- c('lightgrey', 'salmon','red')
featurecolor <- c('lightgrey', '#feeea4','#ee6345', '#a70b44')
# (colours = c("#535ca8", "#99d5a4","#feeea4","#ee6345", "#a70b44"))
# FeaturePlot(allsamplesclean2, features = 'Foxp3',cols = featurecolor)
```

## 15. VlnPlot to check markers 
```{r VlnPlot, message=FALSE, warning=FALSE}
VlnPlot(eyenaive,features=c("Bcl6"), pt.size = 0.1)
VlnPlot(eyenaive,features=c("Tgfbr1"), pt.size = 0)

VlnPlot(eyenaive, features=c("Ccr7"),)
VlnPlot(allsamplesclean2, features=c("Cd28"),)

#group Volin plot 2023.7.16
Vlngene <- c('Il2','Ifng','Il17a','Csf2','Il10','Tgfb1')
Vlngene <- c('Tbx21','Rorc','Foxp3') #Tr1
Vlngene <- c('Nrgn','Pdcd1','Lag3','Icos')
Vlngene <- c('Tgfbr1','Tgfbr2','Tgfbr3')
Vlngene <- c('Ccr7','Aff3','Tcf7','Lef1')
Vlngene <- c('Cblb','Nrgn','Dgkz','Lef1')
Vlngene <- c('Ctla4','Pdcd1','Lag3','Tigit')

Vlngene <- c('Ccr3','Ccr4','Ccr8','Cxcr4',"Il4","Il5","Il9") #Th2 negative

#Th marker genes
Vlngene <- c('Spi1','Il4ra','Ccr10','Il18r1')#No th9
Vlngene <- c("Batf",'Ahr','Ccr4','Ccl7','Il13',"Il22")#No th22
Vlngene <- c('Ccr3','Ccr4','Ccr8','Cxcr4',"Il4","Il5","Il9","Il13") #Th2 negative
Vlngene <- c('Cxcr5','Cd40lg','Cxcl13','Spi1',"Il4ra","Ccl7","Il22") #TFH: CD40 ligand: Tnfsf5,CD154,CXCR4,CXCR5, 

Vlngene <- c('Ctla4','Lag3','Cd200','Pdcd1',"Tigit","Cd28","Icos") #Tim3 "Havcr2","Fas"
Vlngene <- c("Nr4a1","Nr4a2","Nr4a3","Nfatc1","Nfatc2","Nfatc3") 
Vlngene <- c('Izumo1r','Nt5e',"Ndrg1","Irf4") #Izumo1r, FR4, Nt5e, CD73
Vlngene <- c('Mc1r','Mc2r',"Mc3r","Mc4r","Mc5r","") # No a-MSH receptor 
Vlngene <- c("Nr4a1","Nr4a2","Nr4a3","Nfatc1","Nfatc2","Nfatc3","Fos", "Jun") 
Vlngene <- c('Nfatc1','Nfatc2',"Nr4a2",'Rara',"Mafg","Cux1")
Vlngene <- c("Egr2",'Rara',"Nfil3","Klf4","Stat3","Stat5a","Mbd2")

VlnPlot(eyenaive,features =  Vlngene,pt.size=0)
VlnPlot(eyenaive,features =  Vlngene,pt.size=0,ncol=1)+NoLegend() 
VlnPlot(eyenaive,features =  Vlngene,pt.size=0.01,ncol=1)+NoLegend() 
```

## 16. Signature module score 
```{r modulescore, message=FALSE, warning=FALSE}
# Look at the current data
head(allsamples)
# Generate modules (Th1) 
cd_features <- list(c('Ifng', 'Tbx21', 'Cxcr3', 'Nkg7', 'Fasl', 'Il2', 'Tnf', 'Il12rb1', 'Il12rb2'))
# Generate modules (Th2)
cd_features <- list(c('Gata3', 'Il4', 'Il5', 'Il13', 'Il10', 'Areg', 'Maf', 'Il17rb'))
# Generate modules (Th17)
cd_features <- list(c('Rorc', 'Il17a', 'Il17f', 'Il22', 'Ccr6', 'Il23r', 'Tmem176a'))
# Generate modules (Tfh)
cd_features <- list(c('Cxcr5', 'Izumo1r', 'Nt5e', 'Bcl6', 'Slamf6', 'Cd200', 'Ccr7', 'Tcf7'))
#TH1,2,17,Tfh Reference paper: https://doi.org/10.1038/s41590-020-00836-7 Table S2: Curated gene signatures

# Generate modules (TCR)
cd_features <- list(c('Cd28', 'Ctla4', 'Icos', 'Pdcd1', 'Btla','Cd80'))
cd_features <- list(c('Cd28', 'Ctla4', 'Icos', 'Pdcd1','Btla','Cd80','Tnfrsf4','Cd40lg','Itgal','Havcr1','Havcr2'))
cd_features <- list(c('Dusp2', 'Nr4a1', 'Egr1', 'Egr2', 'Myc')) 
#Reference: https://doi.org/10.1073/pnas.1220688110 AND 2021 NI

# Generate modules (CD28 20230608)
cd_features <- list(c('Cd28', 'Ctla4', 'Icos', 'Pdcd1', 'Btla','Cd80'))
#GRAP2,PIK3R5,PRKCA,PRKCB,PRKCQ,RAF1,CTLA4,FOS,GRAP2,ITPR1,ITPR2,MALT1,NFKB1,NFKBIA,PIK3R5,PPP3CA,PRKCQ,REL

FeaturePlot(object = allsamplesclean3, reduction="umap",features = 'CD_Features1')  & scale_colour_gradient (low = "lightgrey", high = "red")

FeaturePlot(object = eyenaive, reduction="umap",features = 'CD_Features1')  & scale_colour_gradient (low = "yellow", high = "red", na.value = "grey50")
FeaturePlot(object = allsamples, reduction="umap",features = 'CD_Features1')  & scale_colour_gradientn(colours = (brewer.pal(n = 8, name = "Pastel2")))
```

## 17. DEGs eye cluster vs naive (1 vs 1) (2023.7.18)
```{r clustercomparisons, message=FALSE, warning=FALSE}
# Test umap plot
DimPlot(eyenaive, reduction = "umap", label = TRUE, repel = TRUE)
DimPlot(allsamplesclean3, reduction = "umap", label = TRUE, repel = TRUE)
###0729
anergy17vsnaive = FindMarkers(eyenaive, ident.1=1, ident.2=2)
write.table(anergy17vsnaive, file=paste(outputpath,"anergy17vsnaive.txt",sep=""), quote=F, sep = "\t")
anergy17vsnaivefiltered=dplyr::filter(anergy17vsnaive, p_val_adj < 0.05 & (avg_log2FC > 1 | avg_log2FC < -1))
nrow(anergy17vsnaivefiltered)#372
nrow(anergy17vsnaive)#2715
write.table(anergy17vsnaivefiltered, file=paste(outputpath,"anergy17vsnaivefiltered.txt",sep=""), quote=F, sep = "\t")

Tregvsnaive = FindMarkers(eyenaive, ident.1=3, ident.2=2)
write.table(Tregvsnaive, file=paste(outputpath,"Tregvsnaive.txt",sep=""), quote=F, sep = "\t")
Tregvsnaivefiltered=dplyr::filter(Tregvsnaive, p_val_adj < 0.05 & (avg_log2FC > 1 | avg_log2FC < -1))
nrow(Tregvsnaivefiltered)# 2873
nrow(Tregvsnaive)#398
write.table(Tregvsnaivefiltered, file=paste(outputpath,"Tregvsnaivefiltered.txt",sep=""), quote=F, sep = "\t")

anergyquietvsnaive = FindMarkers(eyenaive, ident.1=0, ident.2=2)
write.table(anergyquietvsnaive, file=paste(outputpath,"anergyquietvsnaive.txt",sep=""), quote=F, sep = "\t")
anergyquietvsnaivefiltered=dplyr::filter(anergyquietvsnaive, p_val_adj < 0.05 & (avg_log2FC > 1 | avg_log2FC < -1))
nrow(anergyquietvsnaivefiltered)#277
nrow(anergyquietvsnaive)#2413
write.table(anergyquietvsnaivefiltered, file=paste(outputpath,"anergyquietvsnaivefiltered.txt",sep=""), quote=F, sep = "\t")

prolifervsnaive = FindMarkers(eyenaive, ident.1=4, ident.2=2)
write.table(prolifervsnaive, file=paste(outputpath,"prolifervsnaive.txt",sep=""), quote=F, sep = "\t")
prolifervsnaivefiltered=dplyr::filter(prolifervsnaive, p_val_adj < 0.05 & (avg_log2FC > 1 | avg_log2FC < -1))
nrow(prolifervsnaivefiltered)#439
nrow(prolifervsnaive)#3361
write.table(prolifervsnaivefiltered, file=paste(outputpath,"prolifervsnaivefiltered.txt",sep=""), quote=F, sep = "\t")
```

## 18. VolcanoPlot (2023.7.18)
```{r differentialbetweeneveryclusters, message=FALSE, warning=FALSE}
###########################
# To fix pvalue 0 hitting the roof

# 2023.7.18 TH17vsnaive (pre-logfc.threshold = 0.25)
DimPlot(allsamplesclean2, reduction = "umap", label = TRUE, repel = TRUE)
th17vsnaive = FindMarkers(allsamplesclean2, ident.1=2, ident.2=1,logfc.threshold = 0)
pval_0 <- subset(th17vsnaive, th17vsnaive$p_val_adj == 0)
# creating a random list of numbers around 1e-300. 
pval_jitter <- round(rnorm(nrow(pval_0), mean =195, sd = 20),0) 
pval_jitter <- 1*10^-pval_jitter
# replace old p-values by new values
pval_0$p_val_adj <- pval_jitter
# creating a new data set for visualization 
resA_1 <- subset(th17vsnaive,th17vsnaive$p_val_adj!=0)
resA_2 <- rbind(resA_1, pval_0)

# Set custom colors and labels
keyvals <- ifelse(
    resA_2$avg_log2FC < -1, 'dodgerblue3',
      ifelse(resA_2$avg_log2FC > 1, 'brown2',
        'grey'))
  keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == 'brown2'] <- 'Upregulated'
  names(keyvals)[keyvals == 'grey'] <- 'No Change'
  names(keyvals)[keyvals == 'dodgerblue3'] <- 'Downregulated'

  EnhancedVolcano(resA_2 , 
                rownames(resA_2 ),
                x ="avg_log2FC", 
                y ="p_val_adj",
                pCutoff = 10e-10,
                FCcutoff = 1,
                pointSize = 1.0,
                labSize = 0,
                colCustom = keyvals)

head(th17vsanergic)
write.table(th17vsanergic, file=paste(outputpath,"th17vsanergic3all.txt",sep=""), quote=F, sep = "\t")
Pth17vsanergic<-read.delim("/Users/pengz4/Desktop/2023.1 10x data analysis/2023.2.10 EyeCell and Naive output/th17vsanergic3all.txt",sep = "\t")  
  
# 2023.8.8 Tregvsnaive (pre-logfc.threshold = 0.25)
DimPlot(allsamplesclean2, reduction = "umap", label = TRUE, repel = TRUE)
Tregvsnaive = FindMarkers(allsamplesclean2, ident.1=3, ident.2=1,logfc.threshold = 0)
pval_0 <- subset(Tregvsnaive, Tregvsnaive$p_val_adj == 0)
# creating a random list of numbers around 1e-300. 
pval_jitter <- round(rnorm(nrow(pval_0), mean =195, sd = 20),0) 
pval_jitter <- 1*10^-pval_jitter
# replace old p-values by new values
pval_0$p_val_adj <- pval_jitter
# creating a new data set for visualization 
resA_1 <- subset(Tregvsnaive,Tregvsnaive$p_val_adj!=0)
resA_2 <- rbind(resA_1, pval_0)

# Set custom colors and labels
keyvals <- ifelse(
    resA_2$avg_log2FC < -1, 'dodgerblue3',
      ifelse(resA_2$avg_log2FC > 1, 'brown2',
        'grey'))
  keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == 'brown2'] <- 'Upregulated'
  names(keyvals)[keyvals == 'grey'] <- 'No Change'
  names(keyvals)[keyvals == 'dodgerblue3'] <- 'Downregulated'

  EnhancedVolcano(resA_2 , 
                rownames(resA_2 ),
                x ="avg_log2FC", 
                y ="p_val_adj",
                pCutoff = 10e-10,
                FCcutoff = 1,
                pointSize = 1.0,
                labSize = 2,
                colCustom = keyvals)
  
 # 2023.8.8 anergyquietvsnaive (pre-logfc.threshold = 0.25)
DimPlot(allsamplesclean2, reduction = "umap", label = TRUE, repel = TRUE)
anergyquietvsnaive = FindMarkers(allsamplesclean2, ident.1=0, ident.2=1,logfc.threshold = 0)
pval_0 <- subset(anergyquietvsnaive, anergyquietvsnaive$p_val_adj == 0)
# creating a random list of numbers around 1e-300. 
pval_jitter <- round(rnorm(nrow(pval_0), mean =195, sd = 20),0) 
pval_jitter <- 1*10^-pval_jitter
# replace old p-values by new values
pval_0$p_val_adj <- pval_jitter
# creating a new data set for visualization 
resA_1 <- subset(anergyquietvsnaive,anergyquietvsnaive$p_val_adj!=0)
resA_2 <- rbind(resA_1, pval_0)

# Set custom colors and labels
keyvals <- ifelse(
    resA_2$avg_log2FC < -1, 'dodgerblue3',
      ifelse(resA_2$avg_log2FC > 1, 'brown2',
        'grey'))
  keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == 'brown2'] <- 'Upregulated'
  names(keyvals)[keyvals == 'grey'] <- 'No Change'
  names(keyvals)[keyvals == 'dodgerblue3'] <- 'Downregulated'

  EnhancedVolcano(resA_2 , 
                rownames(resA_2 ),
                x ="avg_log2FC", 
                y ="p_val_adj",
                pCutoff = 10e-10,
                FCcutoff = 1,
                pointSize = 1.0,
                labSize = 0,
                colCustom = keyvals)

# Volcano map eye-Treg vs non-Treg (2023.8.8)
DimPlot(allsamplesclean2, reduction = "umap", label = TRUE, repel = TRUE)
TregvsnonTreg = FindMarkers(allsamplesclean2, ident.1=3, ident.2= c(0,2),logfc.threshold = 0)
write.table(TregvsnonTreg, file=paste(outputpath,"TregvsnonTregall.txt",sep=""), quote=F, sep = "\t")
pval_0 <- subset(TregvsnonTreg, TregvsnonTreg$p_val_adj == 0)
  
# creating a random list of numbers around 1e-300. 
pval_jitter <- round(rnorm(nrow(pval_0), mean =195, sd = 20),0) 
pval_jitter <- 1*10^-pval_jitter
# replace old p-values by new values
pval_0$p_val_adj <- pval_jitter
# creating a new data set for visualization 
resA_1 <- subset(TregvsnonTreg,TregvsnonTreg$p_val_adj!=0)
resA_2 <- rbind(resA_1, pval_0)

# Set custom colors and labels
keyvals <- ifelse(
    resA_2$avg_log2FC < -1, 'dodgerblue3',
      ifelse(resA_2$avg_log2FC > 1, 'brown2',
        'grey'))
  keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == 'brown2'] <- 'Upregulated'
  names(keyvals)[keyvals == 'grey'] <- 'No Change'
  names(keyvals)[keyvals == 'dodgerblue3'] <- 'Downregulated'

  EnhancedVolcano(resA_2 , 
                rownames(resA_2 ),
                x ="avg_log2FC", 
                y ="p_val_adj",
                pCutoff = 10e-10,
                FCcutoff = 1,
                pointSize = 1.0,
                labSize = 0,
                colCustom = keyvals)
```

## 19. Dotplot
```{r Dotplot, message=FALSE, warning=FALSE}
dot_markers = c("Grap2","Prkcq","Prkcb","Card11","Pik3cd","Pik3r5","Pdk1","Pag1")
# "Prkcz",<10%,"Stat1","Stat3"

DotPlot(eyenaive, features = dot_markers)+coord_flip()+theme_bw()+#move background 
theme(panel.grid = element_blank(),  
axis.text.x=element_text(angle=90,hjust = 1,vjust=0.5))+#display labels 90  
scale_color_gradientn(values = seq(0,1,0.2),colours = c('#330066','#336699','#66CC66','#FFCC33'))  
labs(x=NULL,y=NULL)+guides(size=guide_legend(order=3)) 

ggsave("dotplot1.pdf", plot = plotdot, width = 5, height = 6)
```

## 21. Monocle2 with R=0.2 (2023.5.24)
```{r monocle2 2023.5.24}
library(monocle)
##subset eye cells.
DimPlot(allsamplesclean2, reduction = "umap", label = T, repel = TRUE) 
eye4subset <- subset(allsamplesclean2, idents = c(0,2,3,4))
DimPlot(eye4subset, reduction = "umap", label = T, repel = TRUE) 
Idents(eye4subset) <- "RNA_snn_res.0.2"

## Assign sample types metadata column 
eye4subset$eye4subset_annotation=recode(as.character(eye4subset$RNA_snn_res.0.2),
 "0"="anergic_quiet","2"="anergic_Il17","3"="Treg", "4"="proliferative") 
Idents(eye4subset) <- "eye4subset_annotation"
levels(eye4subset)

expression_matrix4 = eye4subset@assays$RNA@counts 
cell_metadata4 <- data.frame(group = eye4subset[['orig.ident']],clusters = Idents(eye4subset))
gene_annotation4 <- data.frame(gene_short_name = rownames(expression_matrix4), stringsAsFactors = F) 
rownames(gene_annotation4) <- rownames(expression_matrix4)
pd4 <- new("AnnotatedDataFrame", data = cell_metadata4)
fd4 <- new("AnnotatedDataFrame", data = gene_annotation4)
HSMM4 <- newCellDataSet(expression_matrix4,
                       phenoData = pd4,
                       featureData = fd4,
                       expressionFamily=negbinomial.size())
###change based on https://github.com/cole-trapnell-lab/monocle-release/issues/419
HSMM4 <- detectGenes(HSMM4,min_expr = 0.1)
HSMM_myo4 <- estimateSizeFactors(HSMM4)   #3min
HSMM_myo4 <- estimateDispersions(HSMM_myo4) 

disp_table <- dispersionTable(HSMM_myo4)
disp.genes <- subset(disp_table, mean_expression >= 0.1 )
disp.genes <- as.character(disp.genes$gene_id)

# 2023.5.24 run in Biowulf
head(HSMM4@phenoData@data)
table(HSMM4@phenoData@data$clusters)
#####5.24 
HSMM_myo4 <- reduceDimension(HSMM_myo4, max_components = 2, method = 'DDRTree') 
table(HSMM_myo4@phenoData@data$State)
head(HSMM_myo4@phenoData@data)

HSMM_myo4 <-orderCells(HSMM_myo4,reverse = T)
HSMM_myo4 <-orderCells(HSMM_myo4,root_state = 1)

#### Three Default trajectory plots. Pseudotime, State, Cluster. 
plot1 <-plot_cell_trajectory(HSMM_myo4, color_by = "Pseudotime")
plot2 <- plot_cell_trajectory(HSMM_myo4, color_by = "State")
plot3 <- plot_cell_trajectory(HSMM_myo4, color_by = "clusters")

plot1 <- plot_cell_trajectory(HSMM_myo4, color_by = "Pseudotime")+scale_color_gradientn(colours = c('#330066','#336699','#66CC66','#FFCC33'))
plot3 <- plot_cell_trajectory(HSMM_myo4, color_by = "clusters")+scale_color_manual(values = c("#23acde", "#bfa708","#de6de8", "#22c981"))
ggsave("monocle_eye4time.pdf", plot = plot1, width = 7, height = 7)

### BEAM analysis (long time) 
BEAM_res <- BEAM(HSMM_myo4, branch_point = 1, cores = 20, progenitor_method = "duplicate")
BEAM_res <- BEAM_res[order(BEAM_res$qval),]
BEAM_res <- BEAM_res[,c("gene_short_name", "pval", "qval")]
head(BEAM_res)
mycds_sub_beam <- HSMM_myo4[row.names(subset(BEAM_res, qval < 1e-30)),]

pdf("0615branch1_e50_BEAMk5.pdf", width = 8, height = 80)
plot_genes_branched_heatmap(mycds_sub_beam, branch_point = 1, num_clusters = 5, cores = 20,show_rownames = T)
dev.off()

### BEAM gene list (double check the branch point, k= num_clusters 5 as test, depend on data)
p30 <- plot_genes_branched_heatmap(mycds_sub_beam, branch_point = 1, return_heatmap=T,num_clusters =5, cores = 24,show_rownames = FALSE)
p30$ph$tree_row
clusters <- cutree(p30$ph$tree_row, k = 5)
clustering <- data.frame(clusters)
clustering[,1] <- as.character(clustering[,1])
colnames(clustering) <- "Gene_Clusters"
head(clustering)
write.csv(clustering,'BEAMk5q_30.csv',row.names = T)

##### 2023.5.25 pick the genes in the BEAM List to visualize
# kinetic_genes c("Dgkh", "Pik3r5", "Cblb")))
#c("Nr4a3","Tmem163", "Sell","Rorc", "Ctla4", "Bach2","Il2ra", "Cd28")))
pdf("kinetic0801.pdf", width = 4, height = 6)
kinetic_genes <- row.names(subset(fData(HSMM_myo4),
          gene_short_name %in% c("Foxp3","Tcf7", "Tox2", "Nr4a3")))

plot_genes_branched_pseudotime(HSMM_myo4[kinetic_genes,],
                       branch_point = 1,color_by = "clusters",ncol = 1, cell_size=0.2,branch_labels = c("angery state","Treg state" )) +scale_color_manual(values = c("#23acde","#bfa708","#de6de8", "#22c981"))
dev.off()
```

## 24. SCENIC (after pySCENIC)
```{r scenic, message=FALSE, warning=FALSE}
devtools::install_github('aertslab/SCopeLoomR')
devtools::install_github("aertslab/SCENIC")
BiocManager::install(c("AUCell", "RcisTarget","GENIE3"))
install.packages("optparse")
library(SCopeLoomR)
library(AUCell)
library(SCENIC)
library(dplyr)
library(KernSmooth)
library(plotly)
library(BiocParallel)
library(optparse)
library(grid)
library(data.table)
library(patchwork)
library(stringr)
library(circlize)

#### 1.Read out_SCENIC.loom 
loom <- open_loom("/Users/pengz4/Desktop/2023.1 10x data analysis/2023.2.10 EyeCell and Naive output/pySCENICoutput/out_eyenaiveSCENIC.loom") 
regulons_incidMat <- get_regulons(loom, column.attr.name="Regulons")
regulons_incidMat[1:4,1:4]  
regulons <- regulonsToGeneLists(regulons_incidMat)
regulonAUC <- get_regulons_AUC(loom,column.attr.name='RegulonsAUC')
regulonAucThresholds <- get_regulon_thresholds(loom)
tail(regulonAucThresholds[order(as.numeric(names(regulonAucThresholds)))])
embeddings <- get_embeddings(loom)  
close_loom(loom)

rownames(regulonAUC)
names(regulons)

load("SeuratObjecteyeNaive5.24.RData")
Idents(eyenaive) <- "eyenaive_annotation"
DimPlot(eyenaive,reduction = "umap",label=T ) 
## Assign sample types metadata column 
eyenaive$eyenaive_annotation=recode(as.character(eyenaive$RNA_snn_res.0.2),
                                            "0"="anergic_quiet",
                                            "1"="anergic_IL17",
                                            "2"="naiveT",
                                            "3"="Treg",
                                            "4"="Proliferative")

##### 2. Regulon_scaled Heatmap
sub_regulonAUC <- regulonAUC[,match(colnames(eyenaive),colnames(regulonAUC))]
dim(sub_regulonAUC) #[1]  255 8115

# Check
identical(colnames(sub_regulonAUC), colnames(eyenaive))

cellClusters <- data.frame(row.names = colnames(eyenaive), 
                           seurat_clusters = as.character(eyenaive$eyenaive_annotation))
cellTypes <- data.frame(row.names = colnames(eyenaive), 
                        celltype = eyenaive$eyenaive_annotation)
head(cellTypes)
head(cellClusters)
sub_regulonAUC[1:10,1:10] 
head(sub_regulonAUC)
#Save
save(sub_regulonAUC,cellTypes,cellClusters,eyenaive_annotation,
     file = 'SCENIC_clean3.Rdata')

head(eyenaive)
selectedResolution <- "celltype" # select resolution
cellsPerGroup <- split(rownames(cellTypes), 
                       cellTypes[,selectedResolution])
# remove extened regulons
sub_regulonAUC <- sub_regulonAUC[onlyNonDuplicatedExtended(rownames(sub_regulonAUC)),] 
dim(sub_regulonAUC)  ####[1]  255 8115
# Calculate average expression:
regulonActivity_byGroup <- sapply(cellsPerGroup,
                                  function(cells) 
                                    rowMeans(getAUC(sub_regulonAUC)[,cells]))
# Scale expression. 
regulonActivity_byGroup_Scaled <- t(scale(t(regulonActivity_byGroup),
                                          center = T, scale=T)) 
dim(regulonActivity_byGroup_Scaled)  #[1] 255.   5
regulonActivity_byGroup_Scaled=na.omit(regulonActivity_byGroup_Scaled)
write.csv(regulonActivity_byGroup_Scaled, file = "regulonActivity_byGroup_Scaled_eyenaive0928.csv")

regulonActivity_byGroup_Scaled [1:5,1:5]

### full list heat map of Regulon
Heatmap(
  regulonActivity_byGroup_Scaled,
  name                         = "z-score",
  col                          = colorRamp2(seq(from=-2,to=2,length=11),rev(brewer.pal(11, "RdBu"))),
  show_row_names               = T,
  show_column_names            = TRUE,
  row_names_gp                 = gpar(fontsize = 4),
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  row_title_rot                = 0,
  cluster_rows                 = TRUE,
  cluster_row_slices           = TRUE,
  cluster_columns              = TRUE)

#### Selected Regulon Heatmap
Reguloneye = read.delim("/Users/pengz4/Desktop/2023.8 manuscripts/2023.8 SCENIC/regulonActivity_byGroup_Scaled_eyenaive0928.txt",header = T,sep="\t",row.names = 1)
View(Reguloneye)
Heatmap(
  Reguloneye,
  name                         = "Regulon activity",
  col                          = colorRampPalette(c( "steelblue3","white", "tomato3"))(200),
  show_row_names               = T,
  show_column_names            = TRUE,
  row_names_gp                 = gpar(fontsize = 12),
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  row_title_rot                = 0,
  cluster_rows                 = TRUE,
  cluster_row_slices           = T,
  cluster_columns              = T)

##### 3. visualization
regulonsToPlot = c('Egr1(+)','Egr2(+)','Egr3(+)')
regulonsToPlot %in% row.names(sub_regulonAUC)
eyenaive@meta.data = cbind(eyenaive@meta.data ,t(assay(sub_regulonAUC[regulonsToPlot,])))

# Vis
DotPlot(eyenaive, features = unique(regulonsToPlot)) + RotatedAxis()
RidgePlot(eyenaive, features = regulonsToPlot , ncol = 2) 
VlnPlot(eyenaive, features = regulonsToPlot,pt.size = 0)

FeaturePlot(eyenaive,features = regulonsToPlot)

#2023.9.20
library(scCustomize)
regulonsToPlot = c('Rara(+)','Rxra(+)',"Foxp3(+)","Ikzf2(+)")
regulonsToPlot = c("Egr2(+)","Egr3(+)","Mafg(+)")
#'Nr4a2(+)','Nfatc1(+)',"Egr2(+)",'Rara(+)','Nfatc2(+)',"Mafg(+)","Cux1(+)","Nfil3(+)","Klf4(+)",'Pparg(+)','Bach2(+)',"Batf(+)",'Elf4(+)','Ahr(+)',"Rxra(+)","Prdm1(+)","Elf4(+)","Ikzf2(+)"
regulonsToPlot %in% row.names(sub_regulonAUC)
eyenaive@meta.data = cbind(eyenaive@meta.data ,t(assay(sub_regulonAUC[regulonsToPlot,])))
FeaturePlot_scCustom(eyenaive,features = regulonsToPlot,colors_use = viridis_inferno_light_high)
```


```{r save, message=FALSE, warning=FALSE}
# Save QC session data in the working directory
# Identify the name of the Seurat object you want to delete
seurat_obj_name <- "name_of_seurat_object"
# Remove the Seurat object from your R environment
rm(list = seurat_obj_name)
# Save the modified R environment into a new RData file
save(list = ls(), file = "new_file.RData")

setwd("/Users/pengz4/Desktop/2023.1 10x data analysis/2023.2.10 EyeCell and Naive output/")
save.image(file=paste(outputpath,"SeuratObjecteyeNaive5.24.RData",sep=""))

head(allsamplesclean3)
sessionInfo()
```
