
---
# This RMD file loads the cellranger count data and performs single-cell gene expression analysis of T cell differentiation in an ocular immune privilege model.
---
```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
options(knitr.duplicate.label = "allow")
knitr::opts_chunk$set()
```

## 1. Introduction
This document takes the cellranger count data for each of the samples, does qc anlaysis, filtering, generates before and after qc plots and combines the data.
```{r packages, message=FALSE, warning=FALSE}
install.packages("Seurat")

library(hdf5r)
library(Seurat)
library(ggplot2)
library(cowplot)
library(tidyverse)
library(HGNChelper)
library(EnhancedVolcano)
library(ComplexHeatmap) 
library(patchwork)
```

## 2. Data setup
```{r datasetup, message=FALSE, warning=FALSE}
# Set the input, output paths
directorypath="/path/to/your/project/Results/cellranger/"
outputpath="/path/to/your/project/Ocular_privilege_Output/"
samples=c("S1_HTO_1","S1_HTO_2","S1_HTO_3","S1_HTO_4","S3_HTO_1")
setwd("/path/to/your/project/Ocular_privilege_Output/")
```

## 3. Quality Analysis and Filtering
This part of the workflow reads in each of the samples, analyses the quality, performs the quality filter and merges the data. Both merged unfiltered and merged filtered data are then used for plotting the quality
```{r qc, message=FALSE, warning=FALSE}
# Using the for loop to reads the samples list
# Create empty lists to hold the combined data set
allsamplestomerge=list()
allsamplestomergeprefilter=list()

# Using the for loop to reads the samples list
for(i in samples)
{
  # Remove any old variables
  rm(onesampledataFilt)
  rm(onesampledata)
  rm(onesample)
  # Print the current sample name
  print(i)
  # Read the data file
  onesample <- Read10X_h5(paste0(directorypath,paste0(i,"/sample_filtered_feature_bc_matrix.h5",sep=""),sep=""), use.names = T)

  # Print column names
  colnames(onesample$`Gene Expression`)
  # Add gene names to columnnames
  colnames(onesample$`Gene Expression`) <-     paste(sapply(strsplit(colnames(onesample$`Gene Expression`),split="-"),'[[',1L),i,sep="-")
  # Print column names
  colnames(onesample$`Gene Expression`)
  
  # Create a Seurat data object from the gex matrix, keeping all the features, and only cells with at least 300 genes
  onesampledata <- CreateSeuratObject(counts = onesample[["Gene Expression"]], min.cells = 0,min.features = 300,names.field = 2,names.delim = "\\-")
  
  # Check genes names
  head(rownames(onesampledata))
  # Check data rows and columns
  head(onesampledata)
  
  # Generate Percent mito for each of the cells
  # Lookup genes
  grep("^mt-",rownames(onesampledata@assays$RNA@counts),value = TRUE)
  grep("Rp[sl]",rownames(onesampledata@assays$RNA@counts),value = TRUE)
  grep("^Hb[^(p)]",rownames(onesampledata@assays$RNA@counts),value = TRUE)
  
  # Generate Percent mito for each of the cells
  onesampledata$percent.mito <- PercentageFeatureSet(onesampledata, pattern = "^mt-")
  summary(onesampledata$percent.mito)
  head(onesampledata)
  
  # Generate Percent ribo for each of the cells
  onesampledata$percent.ribo <- PercentageFeatureSet(onesampledata, pattern = "Rp[sl]")
  summary(onesampledata$percent.ribo)
  head(onesampledata)
  
  # Generate Percent hemoglobin for each of the cells
  onesampledata$percent.hb <- PercentageFeatureSet(onesampledata, pattern = "^Hb[^(p)]")
  summary(onesampledata$percent.hb)
  head(onesampledata)
  
  # Check the total number of cells in this sample
  table(onesampledata$orig.ident)
  # Filter data and keep only cells that has percent.mito less than or equal to 8
  onesampledataFilt <- subset(onesampledata, percent.mito <= 8)
  # Check the number of cells after mito filtering
  table(onesampledataFilt$orig.ident)
  # Filter for nCount RNA (total rna molecules sequenced)
  onesampledataFilt <- subset(onesampledataFilt, nCount_RNA >= 1000 & nCount_RNA <= 40000)
  # Check number of cells after ncount filtering
  table(onesampledataFilt$orig.ident)
  # Filter for nfeature (total number of genes expressed)
  onesampledataFilt <- subset(onesampledataFilt, nFeature_RNA >= 400)
  # Check number of cells after nfeature filtering
  table(onesampledataFilt$orig.ident)
  TRUE# Filter for percent ribo
  onesampledataFilt <- subset(onesampledataFilt, percent.ribo > 10 & percent.ribo < 45)
  # Counts after filtering
  table(onesampledataFilt$orig.ident)
  
  # Add filtered and unfiltered data to a list
  allsamplestomerge[[i]]=onesampledataFilt
  allsamplestomergeprefilter[[i]]=onesampledata
}
```
## 4. Merge samples
```{r merge, message=FALSE, warning=FALSE}
# Create one and the rest of the samples as a list, for filtered data
one=allsamplestomerge[[1]]
one
rest=allsamplestomerge[2:length(allsamplestomerge)]
rest

# Create one and the rest of the samples as a list, for profiltered data
onepre=allsamplestomergeprefilter[[1]]
onepre
restpre=allsamplestomergeprefilter[2:length(allsamplestomergeprefilter)]
restpre

# Merge filtered seurat objects
allsamples <- merge(one, y = rest)
allsamples
head(allsamples)
table(Idents(allsamples))

# Merge prefiltered seurat objects
allsamplespre <- merge(onepre, y = restpre)
allsamplespre
head(allsamplespre)
table(Idents(allsamplespre))
```
## 5. Add metadata
```{r metadata, message=FALSE, warning=FALSE}
## Assign sample types metadata column
allsamples$type <- plyr::mapvalues(
  x = allsamples$orig.ident, 
  from = c("S1_HTO_1","S1_HTO_2","S1_HTO_3","S1_HTO_4","S3_HTO_1"),
  to = c("eye","eye","eye","eye","Naive T"))
head(allsamples)

# View type counts
table(allsamples$type)
```
## 6. Generate before and after QC plots
```{r plots, message=FALSE, warning=FALSE}
Idents(allsamples) <- "orig.ident"
# Generate before filtering quality plots
qcbefore=VlnPlot(allsamplespre,features = c("nFeature_RNA", "nCount_RNA","percent.mito","percent.ribo"),ncol = 2, pt.size = 0) +
  NoLegend()
qcbefore
# Save the plot in current working directory
# ggsave(paste(outputpath,"before-qc-violineplot.pdf",sep=""), plot = qcbefore, width = 15, height = 20, units = "cm")

# Generate after filtering quality plots
qcafter=VlnPlot(allsamples,features = c("nFeature_RNA", "nCount_RNA","percent.mito","percent.ribo"),ncol = 2, pt.size = 0) +
  NoLegend()
qcafter
# Save the plot in current working directory
# ggsave(paste(outputpath,"after-qc-violineplot.pdf",sep=""), plot = qcafter, width = 15, height = 20, units = "cm")
```
## 7. Normalization
```{r normalize, message=FALSE, warning=FALSE}
###################### Normalize Data
# Feature counts for each cell are divided by the total counts for that cell and multiplied by the scale.factor. This is then natural-log transformed
# Cell level normalization - accounts for sequencing depth
allsamples <- NormalizeData(
  object = allsamples,
  normalization.method = "LogNormalize",
  scale.factor = 10000)

##################### Find variable genes
# FindVariableFeatures needs normalization
allsamples <- FindVariableFeatures(
  object = allsamples,
  selection.method = "vst")
```
## 8. Scaling & Batch Correction
```{r scale, message=FALSE, warning=FALSE}
##################### SCALE DATA
# PCA needs scaled data
# Zero centers and scales (mean/sd) gene/feature data in each cell (for across sample comparison), so extreme ranges in expression do not affect clustering (done for making cells with similar expression cluster together)
dim(allsamples)
allsamples <- ScaleData(
  object = allsamples,
  # Remove unwanted sources of variation (technical, batch etc.)
  vars.to.regress = c("percent.mito", "nFeature_RNA", "percent.ribo", "nCount_RNA"))
```
## 9. PCA
```{r PCA, message=FALSE, warning=FALSE}
##################### PCA
# Run PCA with 50 pcs
allsamples <- RunPCA(object = allsamples, npcs=50)
# Print genes in the top 5 pcs
print(allsamples[["pca"]], dims = 1:5, nfeatures = 5)
# Identify number of PCs that explains majority of variations
ElbowPlot(allsamples, ndims=50, reduction = "pca")
# Visualize the genes in the first PC
VizDimLoadings(allsamples, dims = 1, ncol = 1) + theme_minimal(base_size = 8)
# Visualize the genes in the second PC
VizDimLoadings(allsamples, dims = 2, ncol = 1) + theme_minimal(base_size = 8)

Idents(allsamples) <- "orig.ident"
# Plot heatmap with cells=500 plotting cells with extreme cells on both ends of spectrum
DimHeatmap(object = allsamples, dims = 1:6, cells = 500, balanced = TRUE)
```

## 10. Clustering
```{r clustering, message=FALSE, warning=FALSE}
##################### Clustering
# use the first 20 pc's
use.pcs = 1:20
# FindNeighbors with 20 pcs
allsamples <- FindNeighbors(allsamples, reduction="pca", dims = use.pcs)
# FindClusters with resolution starting at 0.05 and ending at 0.65, with 0.05 increment
allsamples <- FindClusters(
  object = allsamples,
  resolution = seq(0.05,0.65,0.05),
  verbose = FALSE)
head(allsamples)
# Count number of clusters at each resolution
sapply(grep("res",colnames(allsamples@meta.data),value = TRUE),
       function(x) length(unique(allsamples@meta.data[,x])))
```

## 11. UMAP 
```{r umap, message=FALSE, warning=FALSE}
########################## UMAP
# Run umap with 20 pcs
set.seed(20)
allsamples <- RunUMAP(
  object = allsamples,
  reduction.use = "pca",
  dims = use.pcs)
# Plot umap
Idents(allsamples) <- "RNA_snn_res.0.2"
DimPlot(object = allsamples, pt.size=0.2, reduction = "umap", label = T)
```

### 12. Remove non-T cell:allsamplesclean and re-UMAP
```{r subset, message=FALSE, warning=FALSE}
###########################
# After checking the lineage genes, cluster 5 expresses many myeloid-related genes
# Remove "Macrophages" in r=0.2
Idents(allsamples) <- "RNA_snn_res.0.2"
table(Idents(allsamples))
allsamplesclean <- subset(allsamples, idents = 5 , invert = TRUE)
table(Idents(allsamplesclean))
DimPlot(allsamplesclean, reduction = "umap", label = T, repel = TRUE) 

# Assign "eyenaive" object for downstream analysis
eyenaive<-allsamplesclean
# re-PCA.
# Run PCA with 50 pcs
eyenaive <- RunPCA(object = eyenaive, npcs=50)
# Print genes in the top 5 pcs
print(eyenaive[["pca"]], dims = 1:5, nfeatures = 5)
# Identify number of PCs that explains majority of variations
ElbowPlot(eyenaive, ndims=50, reduction = "pca")
# Visualize the genes in the first PC
VizDimLoadings(eyenaive, dims = 1, ncol = 1) + theme_minimal(base_size = 8)
# Visualize the genes in the second PC
VizDimLoadings(eyenaive, dims = 2, ncol = 1) + theme_minimal(base_size = 8)

### Re-cluster
eyenaive <- FindNeighbors(eyenaive, reduction="pca", dims = use.pcs)
eyenaive <- FindClusters(
  object = eyenaive,
  resolution = seq(0.1,0.8,0.1),
  verbose = FALSE)
#can try different resolution range and increase degree, e.g., 0.1 or 0.05 

# Count number of clusters at each resolution
sapply(grep("res",colnames(eyenaive@meta.data),value = TRUE),
       function(x) length(unique(eyenaive@meta.data[,x])))

# Re-Run umap with 20 pcs
set.seed(20)
eyenaive <- RunUMAP(
  object = eyenaive,
  reduction.use = "pca",
  dims = use.pcs)

Idents(eyenaive) <- "RNA_snn_res.0.2"
DimPlot(eyenaive, reduction = "umap", label = TRUE, repel = TRUE)
DimPlot(object = eyenaive, pt.size=0.2, reduction = "umap", group.by = "type",label = T)
DimPlot(object = eyenaive, pt.size=0.2, reduction = "umap", split.by = "orig.ident")

# list cell number in each cluster for sample types
table(Idents(eyenaive),eyenaive$type)
table(Idents(eyenaive),eyenaive$orig.ident)
propotioninfo=table(Idents(eyenaive),eyenaive$orig.ident)

# Save the file of cell number in each cluster text file
write.table(propotioninfo, file=paste(outputpath,"propotioninfo.txt",sep=""), quote=F, sep = "\t")


# Show sorted naive before injection and retrieved from eye after on week
DimPlot(object = eyenaive, pt.size=0.2, reduction = "umap", group.by = "type",cols = c("#2dc2a9","salmon"))

Idents(eyenaive) <- "RNA_snn_res.0.2"
DimPlot(object = eyenaive, pt.size=0.2, reduction = "umap")
## Assign sample types metadata column 
#### different version of annotation during analysis. 
#### anergic_naive=nfc1=anergic1, and anergic_IL17=nfc2=anergic2
eyenaive$eyenaive_annotation=recode(as.character(eyenaive$RNA_snn_res.0.2),
                              "0"="anergic_naive",
                              "1"="anergic_IL17",
                              "2"="naiveT",
                              "3"="Treg",
                              "4"="proliferative")

head(eyenaive)
Idents(eyenaive) <- "eyenaive_annotation"
levels(eyenaive)
DimPlot(object = eyenaive, pt.size=0.2, reduction = "umap",label = T)
```

### 13. FindAllMarkers (Suppl Fig. 1A)
```{r FindAllMarkers, message=FALSE, warning=FALSE}
###########################
markersall_eyenaive=FindAllMarkers(eyenaive)
head(markersall_eyenaive)

write.table(markersall_eyenaive, file=paste(outputpath,"markersall_eyenaive.txt",sep=""), quote=F, row.names = FALSE, sep = "\t")
markersall_eyenaivefiltered=dplyr::filter(markersall_eyenaive, p_val_adj < 0.05 & (avg_log2FC > 1 | avg_log2FC < -1))
nrow(markersall_eyenaive)
nrow(markersall_eyenaivefiltered)
write.table(markersall_eyenaivefiltered, file=paste(outputpath,"markersall_eyenaivefiltered.txt",sep=""), quote=F, row.names = FALSE, sep = "\t")

# Extract the top 10 marker for each cluster and plot heatmap 
markersall_eyenaive %>%
  group_by(cluster) %>%
  top_n(n = 10, wt = avg_log2FC) -> top10
top10

# Get only the marker - gene names
top10_eyenaive<- top10 %>%
  select(gene) %>% 
  unclass() %>% 
  stack() %>%
  pull(values) %>%
  unique() %>%
  .[!is.na(.)]
top10_eyenaive

# Scale all genes - not just variable genes 
eyenaiveheatmap = eyenaive
eyenaiveheatmap <- ScaleData(
  object = eyenaive,
features = rownames(eyenaive),
vars.to.regress = c("percent.mito", "nFeature_RNA", "percent.ribo", "nCount_RNA"))

# heatmap Fig.S1
DoHeatmap(eyenaiveheatmap, features = top10_eyenaive, group.by = "RNA_snn_res.0.2") + scale_fill_gradientn(colors = c("steelblue4", "white", "firebrick3"))+  NoLegend()
```

### 14. FeaturePlot to check markers (Fig. 1E, Fig. 2B-C, and Fig.6D)
```{r FeaturePlot, message=FALSE, warning=FALSE}
########################### tables
# Fig.1E Primed phenotype and Treg and proliferation marker
FeaturePlot(eyenaive, features=c("Cd44"),)
FeaturePlot(eyenaive, features=c("Sell"),)
FeaturePlot(eyenaive, features=c("Foxp3"),)
FeaturePlot(eyenaive, features=c("Mki67"),)

# Fig.2B Th lineage-specific markers
FeaturePlot(eyenaive, features=c("Tbx21"),)
FeaturePlot(eyenaive, features=c("Ccr5"),)
FeaturePlot(eyenaive, features=c("Cxcr3"),)

FeaturePlot(eyenaive, features=c("Rorc"),)
FeaturePlot(eyenaive, features=c("Il17a"),)
FeaturePlot(eyenaive, features=c("Ccr6"),)

FeaturePlot(eyenaive, features=c("Il2ra"),)
FeaturePlot(eyenaive, features=c("Il10"),)
FeaturePlot(eyenaive, features=c("Tgfb1"),)

# Fig.2C naive related marker
FeaturePlot(eyenaive, features=c("Ccr7"),) 
FeaturePlot(eyenaive, features=c("Lef1"),) 
FeaturePlot(eyenaive, features=c("Tcf7"),) 

# Fig.6D Reciprocal expression of Foxp3 and Nrgn
FeaturePlot(eyenaive, features=c("Nrgn"),) 

```

### 15. VlnPlot to check markers (Fig. 2A and Suppl. Fig.1 B-C )
```{r VlnPlot, message=FALSE, warning=FALSE}
Idents(eyenaive) <- "RNA_snn_res.0.2"
DimPlot(eyenaive, pt.size=0.2, reduction = "umap")

Idents(eyenaive) <- factor(Idents(eyenaive),levels=c(2,0,1,3,4))

Vlngene <- c('Foxp3', 'Tbx21','Rorc', 'Gata3', 'Bcl6')
Vlngene <- c('Il2', 'Ifng', 'Il17a', 'Csf2', 'Il10', 'Tgfb1')
VlnPlot(eyenaive,features =  Vlngene,pt.size=0,ncol=1)+NoLegend()

#Suppl Fig.1B-C
Vlngene <- c('Il17f','Il22',"Il23r")  # Th17 associated
Vlngene <- c('Ccr3','Ccr8',"Il4", 'Il5') # Th2 associated
Vlngene <- c('Cxcr5','Cxcl13',"Il21") # Tfh associated
Vlngene <- c('Ahr','Ccr4',"Ccl7",'Il13','Il22') # Th22 associated
Vlngene <- c('Spi1','Il4ra',"Il9") # Th9 associated

#Tr1 Reference paper: https://doi.org/10.1016/j.immuni.2018.12.001 Immunity
Vlngene <- c('Itga2','Il10')  # Tr1 associated gene

Vlngene <- c('Izumo1r','Nt5e') # Anergy marker:#FR4,#CD73
Vlngene <- c('Il12a','Ebi3') # Il35

```

### 16. DEGs eye cluster vs naive baseline. VolcanoPlot (For Fig.3, see also Suppl table. 2)
```{r clustercomparisons, message=FALSE, warning=FALSE}
# Test umap plot
DimPlot(eyenaive, reduction = "umap", label = TRUE, repel = TRUE)

nfc1vsnaive = FindMarkers(eyenaive, ident.1=0, ident.2=2)
write.table(nfc1vsnaive, file=paste(outputpath,"DEG_nfc1vsnaive.txt",sep=""), quote=F, sep = "\t")
nfc1vsnaivefiltered=dplyr::filter(nfc1vsnaive, p_val_adj < 0.05 & (avg_log2FC > 1 | avg_log2FC < -1))
nrow(nfc1vsnaivefiltered)
nrow(nfc1vsnaive)
write.table(nfc1vsnaivefiltered, file=paste(outputpath,"DEG_nfc1vsnaivefiltered.txt",sep=""), quote=F, sep = "\t")

nfc2vsnaive = FindMarkers(eyenaive, ident.1=1, ident.2=2)
write.table(nfc2vsnaive, file=paste(outputpath,"DEG_nfc2vsnaive.txt",sep=""), quote=F, sep = "\t")
nfc2vsnaivefiltered=dplyr::filter(nfc2vsnaive, p_val_adj < 0.05 & (avg_log2FC > 1 | avg_log2FC < -1))
write.table(nfc2vsnaivefiltered, file=paste(outputpath,"DEG_nfc2vsnaivefiltered.txt",sep=""), quote=F, sep = "\t")

Tregvsnaive = FindMarkers(eyenaive, ident.1=3, ident.2=2)
write.table(Tregvsnaive, file=paste(outputpath,"DEG_Tregvsnaive.txt",sep=""), quote=F, sep = "\t")
Tregvsnaivefiltered=dplyr::filter(Tregvsnaive, p_val_adj < 0.05 & (avg_log2FC > 1 | avg_log2FC < -1))
write.table(Tregvsnaivefiltered, file=paste(outputpath,"DEG_Tregvsnaivefiltered.txt",sep=""), quote=F, sep = "\t")

prolifervsnaive = FindMarkers(eyenaive, ident.1=4, ident.2=2)
write.table(prolifervsnaive, file=paste(outputpath,"DEG_prolifervsnaive.txt",sep=""), quote=F, sep = "\t")
prolifervsnaivefiltered=dplyr::filter(prolifervsnaive, p_val_adj < 0.05 & (avg_log2FC > 1 | avg_log2FC < -1))
write.table(prolifervsnaivefiltered, file=paste(outputpath,"DEG_prolifervsnaivefiltered.txt",sep=""), quote=F, sep = "\t")

########################### 
# Volcano plot
# nfc1 vs naive
Vnfc1vsnaive<-read.delim("DEG_nfc1vsnaive.txt",sep = "\t")
pval_0 <- subset(Vnfc1vsnaive, Vnfc1vsnaive$p_val_adj == 0)
# creating a random list of numbers around 1e-300. 
pval_jitter <- round(rnorm(nrow(pval_0), mean =150, sd = 35),0) 
pval_jitter <- 1*10^-pval_jitter
# replace old p-values by new values
pval_0$p_val_adj <- pval_jitter
# creating a new data set for visualization 
Vnfc1vsnaive <- subset(Vnfc1vsnaive,Vnfc1vsnaive$p_val_adj!=0)
Vnfc1vsnaive <- rbind(Vnfc1vsnaive, pval_0)
# Set custom colors and labels
keyvals <- ifelse(
  Vnfc1vsnaive$avg_log2FC < -1, 'dodgerblue3',
  ifelse(Vnfc1vsnaive$avg_log2FC > 1, 'brown2',
         'grey'))
keyvals[is.na(keyvals)] <- 'grey'
names(keyvals)[keyvals == 'brown2'] <- 'Upregulated'
names(keyvals)[keyvals == 'grey'] <- 'Non-DEG'
names(keyvals)[keyvals == 'dodgerblue3'] <- 'Downregulated'
p1 <- EnhancedVolcano(Vnfc1vsnaive,rownames(Vnfc1vsnaive),
                      x ="avg_log2FC", y ="p_val_adj",title = 'nfc1 vs naive',
                      selectLab = c("Nrgn","Cblb","Nt5e","Nr4a1","Nr4a2","Nr4a3","Ctla4","Stat3","Tox",
                         "Tgfb1","Maf","Lag3","Ikzf2","Foxp1","Sell","Pik3r5","Satb1","Pag1","S100a4","Cd200",
                         "Nrp1","Tnfrsf4","Pdcd1","Ccl5","Tnfrsf1b", "Tcf7","Itgav","Itga4","Ccr7","Aff3","Il7r","Tox2"),
             pCutoff = 10e-10,FCcutoff = 1,cutoffLineCol = 'darkgrey',vlineCol = 'darkgrey', pointSize = 1.0,
             labSize = 3,vline= c(-0.25,0.25),
             colCustom = keyvals,drawConnectors = TRUE,arrowheads = FALSE,
             lengthConnectors = unit(3, 'npc'),legendIconSize = 0, widthConnectors = 0.5,max.overlaps=50)
p2 <- p1 + ggplot2::coord_cartesian(xlim=c(-3, 3.5)) +
  ggplot2::scale_x_continuous(breaks=seq(-3, 3.5,1))+
  ggplot2::theme_bw()
p2
ggsave("fig3a-nfc1.pdf", plot = p2, width = 6, height = 4)

```

#### 17. SCENIC (pySCENIC)
```{python pySCENIC, message=FALSE, warning=FALSE}
### Prepare and load database for SCENIC
mm10_10kbp_up_10kbp_down_full_tx_v10_clust.genes_vs_motifs.rankings.feather
wget -c https://resources.aertslab.org/cistarget/motif2tf/motifs-v10nr_clust-nr.mgi-m0.001-o0.0.tbl
wget -c https://resources.aertslab.org/cistarget/tf_lists/allTFs_mm.txt

##### Export matrix data from seurat object in R 
write.csv(t(as.matrix(eyenaive@assays$RNA@counts)),file = "eyenaiveforSCENIC.csv")

python3 -m pip install loompy
python3 -m pip install numpy
python3 -m pip install scanpy

# python script, named as change.py
cat >change.py
#creat a new change.py
import os,sys
os.getcwd()
os.listdir(os.getcwd()) 
import loompy as lp;
import numpy as np;
import scanpy as sc;
x=sc.read_csv("eyenaiveforSCENIC.csv");
row_attrs = {"Gene": np.array(x.var_names),};
col_attrs = {"CellID": np.array(x.obs_names)};
lp.create("eyenaivescenic.loom",x.X.transpose(),row_attrs,col_attrs);

#Back to file, execute the change.py
python3 change.py

## execute pySCENIC
conda activate pyscenic

# mice database #chaneg the path accordingly
dir=/gpfs/gsfs10/users/CaspiRNAseqData/Results/SCENIC-figures/
tfs=$dir/allTFs_mm.txt
feather=$dir/mm10_10kbp_up_10kbp_down_full_tx_v10_clust.genes_vs_motifs.rankings.feather
tbl=$dir/motifs-v10nr_clust-nr.mgi-m0.001-o0.0.tbl

input_loom=./eyenaivescenic.loom
ls $tfs  $feather  $tbl  

#2.1 grn
pyscenic grn \
--num_workers 10 \
--output grn.eyenaivescenic.tsv \
--method grnboost2 \
eyenaivescenic.loom \
$tfs 

#2.2 cistarget
pyscenic ctx \
grn.eyenaivescenic.tsv $feather \
--annotations_fname $tbl \
--expression_mtx_fname $input_loom  \
--mode "dask_multiprocessing" \
--output ciseyenaivescenic.csv \
--num_workers 12  \
--mask_dropouts

#2.3 AUCell
pyscenic aucell \
$input_loom \
ciseyenaivescenic.csv \
--output out_eyenaiveSCENIC.loom \
--num_workers 10

#(This step will create an ~11M out_SCENIC.loom file)
# Next section is using this loom file as input for visualization
```

#### 17. SCENIC (visualize after pySCENIC in R)
```{r scenic, message=FALSE, warning=FALSE}
devtools::install_github('aertslab/SCopeLoomR')
devtools::install_github("aertslab/SCENIC")
BiocManager::install(c("AUCell", "RcisTarget","GENIE3"))
library(SCopeLoomR)
library(AUCell)
library(SCENIC)
library(dplyr)
library(KernSmooth)
library(plotly)
library(BiocParallel)
library(optparse)
library(grid)
library(data.table)
library(patchwork)
library(stringr)
library(circlize)

#### 1.Read out_SCENIC.loom 
loom <- open_loom("/path/to/your/project/pySCENICoutput/out_eyenaiveSCENIC.loom") 
regulons_incidMat <- get_regulons(loom, column.attr.name="Regulons")
regulons_incidMat[1:4,1:4]  
regulons <- regulonsToGeneLists(regulons_incidMat)
regulonAUC <- get_regulons_AUC(loom,column.attr.name='RegulonsAUC')
regulonAucThresholds <- get_regulon_thresholds(loom)
tail(regulonAucThresholds[order(as.numeric(names(regulonAucThresholds)))])
embeddings <- get_embeddings(loom)  
close_loom(loom)

rownames(regulonAUC)
names(regulons)

Idents(eyenaive) <- "eyenaive_annotation"
DimPlot(eyenaive,reduction = "umap",label=T ) 
## Assign sample types metadata column 
eyenaive$eyenaive_annotation=recode(as.character(eyenaive$RNA_snn_res.0.2),
                                            "0"="anergic1",
                                            "1"="anergic2",
                                            "2"="naiveT",
                                            "3"="Treg",
                                            "4"="Proliferative")

##### 2. Regulon_scaled Heatmap
sub_regulonAUC <- regulonAUC[,match(colnames(eyenaive),colnames(regulonAUC))]
dim(sub_regulonAUC) #[1]  255 8115

# Check
identical(colnames(sub_regulonAUC), colnames(eyenaive))

cellClusters <- data.frame(row.names = colnames(eyenaive), 
                           seurat_clusters = as.character(eyenaive$eyenaive_annotation))
cellTypes <- data.frame(row.names = colnames(eyenaive), 
                        celltype = eyenaive$eyenaive_annotation)
head(cellTypes)
head(cellClusters)
sub_regulonAUC[1:10,1:10] 
head(sub_regulonAUC)

head(eyenaive)
selectedResolution <- "celltype" # select resolution
cellsPerGroup <- split(rownames(cellTypes), 
                       cellTypes[,selectedResolution])
# remove extened regulons
sub_regulonAUC <- sub_regulonAUC[onlyNonDuplicatedExtended(rownames(sub_regulonAUC)),] 
dim(sub_regulonAUC)  ####[1]  255 8115
# Calculate average expression:
regulonActivity_byGroup <- sapply(cellsPerGroup,
                                  function(cells) 
                                    rowMeans(getAUC(sub_regulonAUC)[,cells]))
# Scale expression. 
regulonActivity_byGroup_Scaled <- t(scale(t(regulonActivity_byGroup),
                                          center = T, scale=T)) 
dim(regulonActivity_byGroup_Scaled)  #[1] 255.   5
regulonActivity_byGroup_Scaled=na.omit(regulonActivity_byGroup_Scaled)
write.csv(regulonActivity_byGroup_Scaled, file = "regulonActivity_byGroup_Scaled_eyenaive.csv")

regulonActivity_byGroup_Scaled [1:5,1:5]

### full list heat map of Regulon
Heatmap(
  regulonActivity_byGroup_Scaled,
  name                         = "z-score",
  col                          = colorRamp2(seq(from=-2,to=2,length=11),rev(brewer.pal(11, "RdBu"))),
  show_row_names               = T,
  show_column_names            = TRUE,
  row_names_gp                 = gpar(fontsize = 4),
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  row_title_rot                = 0,
  cluster_rows                 = TRUE,
  cluster_row_slices           = TRUE,
  cluster_columns              = TRUE)

#### Selected Regulon Heatmap (Fig. 6B)
Reguloneye = read.delim("/path/to/your/project/regulonActivity_byGroup_Scaled_eyenaive0928.txt",header = T,sep="\t",row.names = 1)
View(Reguloneye)
Heatmap(
  Reguloneye,
  name                         = "Regulon activity",
  col                          = colorRampPalette(c( "steelblue3","white", "tomato3"))(200),
  show_row_names               = T,
  show_column_names            = TRUE,
  row_names_gp                 = gpar(fontsize = 12),
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  row_title_rot                = 0,
  cluster_rows                 = TRUE,
  cluster_row_slices           = T,
  cluster_columns              = T)

##### 3. visualization in UMAP
regulonsToPlot = c('Egr1(+)','Egr2(+)','Egr3(+)')
regulonsToPlot %in% row.names(sub_regulonAUC)
eyenaive@meta.data = cbind(eyenaive@meta.data ,t(assay(sub_regulonAUC[regulonsToPlot,])))

DotPlot(eyenaive, features = unique(regulonsToPlot)) 
library(scCustomize)
FeaturePlot_scCustom(eyenaive,features = regulonsToPlot,colors_use = viridis_inferno_light_high)

```

#### 18. Monocle2 with R=0.2 
```{r monocle2, message=FALSE, warning=FALSE}
library(monocle)
##subset eye cells.
Idents(eyenaive) <- "RNA_snn_res.0.2"
DimPlot(eyenaive,reduction = "umap",label=T ) 
eyesubset <- subset(eyenaive, idents = c(0,1,3,4))

DimPlot(eyesubset, reduction = "umap", label = T, repel = TRUE) 
Idents(eyesubset) <- "RNA_snn_res.0.2"

## Assign sample types metadata column 
eyesubset$eye4subset_annotation=recode(as.character(eyesubset$RNA_snn_res.0.2),
 "0"="anergic_quiet","1"="anergic_Il17","3"="Treg", "4"="proliferative") 
Idents(eyesubset) <- "eye4subset_annotation"
levels(eyesubset)

expression_matrix4 = eyesubset@assays$RNA@counts 
cell_metadata4 <- data.frame(group = eyesubset[['orig.ident']],clusters = Idents(eyesubset))
gene_annotation4 <- data.frame(gene_short_name = rownames(expression_matrix4), stringsAsFactors = F) 
rownames(gene_annotation4) <- rownames(expression_matrix4)
pd4 <- new("AnnotatedDataFrame", data = cell_metadata4)
fd4 <- new("AnnotatedDataFrame", data = gene_annotation4)
HSMM4 <- newCellDataSet(expression_matrix4,
                       phenoData = pd4,
                       featureData = fd4,
                       expressionFamily=negbinomial.size())
###change based on https://github.com/cole-trapnell-lab/monocle-release/issues/419
HSMM4 <- detectGenes(HSMM4,min_expr = 0.1)
HSMM_myo4 <- estimateSizeFactors(HSMM4)   
HSMM_myo4 <- estimateDispersions(HSMM_myo4) 

disp_table <- dispersionTable(HSMM_myo4)
disp.genes <- subset(disp_table, mean_expression >= 0.1 )
disp.genes <- as.character(disp.genes$gene_id)

# run in Biowulf
head(HSMM4@phenoData@data)
table(HSMM4@phenoData@data$clusters)
###### DDRTree take time 
HSMM_myo4 <- reduceDimension(HSMM_myo4, max_components = 2, method = 'DDRTree') 
table(HSMM_myo4@phenoData@data$State)
head(HSMM_myo4@phenoData@data)

HSMM_myo4 <-orderCells(HSMM_myo4,reverse = T)

#### Three Default trajectory plots. Pseudotime, State, Cluster. 
plot1 <-plot_cell_trajectory(HSMM_myo4, color_by = "Pseudotime")
plot2 <- plot_cell_trajectory(HSMM_myo4, color_by = "State")
plot3 <- plot_cell_trajectory(HSMM_myo4, color_by = "clusters")

plot_cell_trajectory(HSMM_myo4, color_by = "Pseudotime")
plot1 <- plot_cell_trajectory(HSMM_myo4, color_by = "Pseudotime")+scale_color_gradientn(colours = c('#330066','#336699','#66CC66','#FFCC33'))
plot3 <- plot_cell_trajectory(HSMM_myo4, color_by = "clusters")+scale_color_manual(values = c("#23acde", "#bfa708","#de6de8", "#22c981"))
ggsave("monocle_eye4time.pdf", plot = plot1, width = 7, height = 7)

### BEAM analysis (long time) 
BEAM_res <- BEAM(HSMM_myo4, branch_point = 1, cores = 20, progenitor_method = "duplicate")
BEAM_res <- BEAM_res[order(BEAM_res$qval),]
BEAM_res <- BEAM_res[,c("gene_short_name", "pval", "qval")]
head(BEAM_res)
mycds_sub_beam <- HSMM_myo4[row.names(subset(BEAM_res, qval < 1e-30)),]

plot_genes_branched_heatmap(mycds_sub_beam, branch_point = 1, num_clusters = 5, cores = 20,show_rownames = T)

### BEAM gene list (double check the branch point, k= num_clusters 5 as test, depend on data)
p30 <- plot_genes_branched_heatmap(mycds_sub_beam, branch_point = 1, return_heatmap=T,num_clusters =5, cores = 24,show_rownames = FALSE)
p30$ph$tree_row
clusters <- cutree(p30$ph$tree_row, k = 5)
clustering <- data.frame(clusters)
clustering[,1] <- as.character(clustering[,1])
colnames(clustering) <- "Gene_Clusters"
head(clustering)
write.csv(clustering,'BEAMk5q_30.csv',row.names = T)

##### Selected genes in the BEAM List to visualize 
install.packages("igraph")

pdf("kinetic_bidirection.pdf", width = 4, height = 6)
kinetic_genes <- row.names(subset(fData(HSMM_myo4),
          gene_short_name %in% c("Foxp3","Tcf7", "Tox2", "Nr4a3")))
plot_genes_branched_pseudotime(HSMM_myo4[kinetic_genes,],
                       branch_point = 1,color_by = "clusters",ncol = 1, cell_size=0.2,branch_labels = c("angery state","Treg state" )) +scale_color_manual(values = c("#23acde","#bfa708","#de6de8", "#22c981"))
dev.off()
```

##### 19. RNA velocity analysis (integrate velocity vectors into Seurat object)
```{r Velocyto R, message=FALSE, warning=FALSE}
# run from the file "smb://hpcdrive.nih.gov/CaspiRNAseqData/velocyto/sample1_hto1" ,etc.
install_github("velocyto-team/velocyto.R")
library(velocyto.R)
library(Seurat)
library(SeuratWrappers)

# Samples list
samples=c("S1_HTO_1","S1_HTO_2","S1_HTO_3","S1_HTO_4","S3_HTO_1")
# Read velocity files
allsamplesvelocyto=list()
# Run for each sample separately, change for range accordingly
for(i in 1:4)
{
  print(i)
  sampleid=paste("S1_HTO_",i,sep="")
  print(sampleid)
  filename=paste("/data/CaspiRNAseqData/velocyto/sample1_hto",i,"/sample_alignments.loom",sep="")
  print(filename)
  allsamplesvelocyto[[sampleid]] <- ReadVelocity(file = filename)
}

sampleid=paste("S3_HTO_1")
print(sampleid)
filename=paste("/data/CaspiRNAseqData/velocyto/sample3_hto1","/sample_alignments.loom",sep="")
print(filename)
allsamplesvelocyto[[sampleid]] <- ReadVelocity(file = filename)

# Check to see if all samples are in
names(allsamplesvelocyto)  #### Yes

# create a new allsamplesS variable
allsamplesS = list()
# For renaming cells in seurat
for(j in samples)
{
  # print current sample name
  print(j)
  toappend=paste("-",j,sep="")
  for (i in names(x = allsamplesvelocyto[[j]])) 
  {
    ### Store assay in a new variable
    assay <- allsamplesvelocyto[[j]][[i]]
    ### Rename cell names in loom file to match cell names in Seurat object
    #colnames(assay) <- gsub('Part of cell name to change', 'Changed part', colnames(assay))
    colnames(assay) <- gsub("^sample_alignments_(.*?):","",colnames(assay))
    colnames(assay) <- gsub("x",toappend,colnames(assay))
    print(head(colnames(assay)))
    
    ### Integrate velocyto assay change to original
    allsamplesvelocyto[[j]][[i]] = assay
  }
  allsamplesS[[j]]=as.Seurat(allsamplesvelocyto[[j]])
  allsamplesS[[j]]$orig.ident = j
  print(allsamplesS[[j]])
}

#allsamplevelocyto - original velocyto output with original names
#allsampleS - velocyto converted to seurat with new cell names
#s1hto1 - velocyto converted to seurat for one sample
#allsamplesSall - seurat merged velocyto sample runs
#allsamplesV - seurat object with GEX, original, without velocity

# Merge velocyto only seurat samples for velocyto only velocity analysis
# assign one sample to a variable 
names(allsamplesS) #[1] "S1_HTO_1" "S1_HTO_2" "S1_HTO_3" "S1_HTO_4" "S3_HTO_1"
s3hto1=allsamplesS$S3_HTO_1
# remove that one sample from the group
allsamplesS$S3_HTO_1 = NULL
# check samples
names(allsamplesS)
# merge all samples
allsamplesSall = merge(s3hto1, y=allsamplesS)   
names(allsamplesSall) #[1] "spliced"   "unspliced" "ambiguous"
head(allsamplesSall)

# Normalize
allsamplesSall = NormalizeData(allsamplesSall)
# scale
allsamplesSall = ScaleData(allsamplesSall)
# PCA
allsamplesSall = RunPCA(allsamplesSall, features = rownames(allsamplesSall))
# Findneighbors
allsamplesSall <- FindNeighbors(allsamplesSall, dims = 1:50)
# Find clusters
allsamplesSall <- FindClusters(allsamplesSall)
# umap
allsamplesSall <- RunUMAP(allsamplesSall, dims = 1:50)
# Use previous gex general workflow prepped seurat data
# Assign the gex seurat object from the general gex workflow output

names(allsamplesSall) ## "spliced","unspliced","ambiguous","spliced_nn","spliced_snn", "pca","umap" 
head(allsamplesSall)

allsamplesV = allsamples  
# how many cells match between gex and vel
length(which(is.element(colnames(allsamplesSall), colnames(allsamplesV))))  ####[1] 8149
# subset gex cells from vel
vel <- allsamplesSall[rownames(allsamplesV), colnames(allsamplesV)]
# gex integration
vel_integrated <- vel
# nearest neighbor graph
vel_integrated@graphs[["RNA_nn"]] <- allsamplesV@graphs[["RNA_nn"]][colnames(vel),colnames(vel)]
# cluster assignments
vel_integrated <- AddMetaData(object = vel_integrated,
                              metadata = FetchData(allsamplesV,vars = "RNA_snn_res.0.2",cells = colnames(vel)),
                              col.name = "gex.clusters")
# assign cluster identity
Idents(vel_integrated) <- "gex.clusters"
levels(vel_integrated)     #[1] "0" "1" "2" "3" "4" "5"

######## subset eye-primed clusters
vel_integrated_eye <- subset(x = vel_integrated, idents = c("0","2","3","4"))
# format clusters as factors
vel_integrated_eye$gex.clusters = factor(vel_integrated_eye$gex.clusters)
# run umap. 
vel_integrated_eye <- RunUMAP(vel_integrated_eye, dims = 1:50)
# velocity using original gex, run velocity, using default parameters, with 60 cores
# please refer https://rdrr.io/github/satijalab/seurat-wrappers/man/RunVelocity.html
vel_integrated_eyeV <- RunVelocity(vel_integrated_eye,    
                                   deltaT = 1,
                                   kCells = 25,
                                   fit.quantile = 0.02,
                                   verbose = FALSE,
                                   ncores = 60)
# set colors and name cluster colors
fourcolors <- c("#bfa708","#48c785","#33a7f5", "#d272ed")
ident_colors <- fourcolors
names(ident_colors) <- levels(vel_integrated_eyeV$gex.clusters)
cell_colors <- ident_colors[vel_integrated_eyeV$gex.clusters]
names(cell_colors) <- colnames(vel_integrated_eyeV)

# generate velocity plot  Takes time
# Draw velocity using its own umap data and standard parameters
# For parameter details refer https://rdrr.io/github/velocyto-team/velocyto.R/man/show.velocity.on.embedding.cor.html
pdf("velocity_eye_plot.pdf")
show.velocity.on.embedding.cor(emb = Embeddings(vel_integrated_eyeV,reduction = "umap"),
                               vel = Tool(vel_integrated_eyeV,slot = "RunVelocity"),
                               n = 200,
                               scale = "sqrt",
                               xlab = colnames(Embeddings(vel_integrated_eyeV,reduction = "umap"))[1],
                               ylab = colnames(Embeddings(vel_integrated_eyeV,reduction = "umap"))[2],
                               cell.colors = ac(x = cell_colors, alpha = 0.5),
                               cex = 0.8,
                               arrow.scale = 3,
                               show.grid.flow = TRUE,
                               min.grid.cell.mass = 0.5,
                               grid.n = 40,
                               arrow.lwd = 1,
                               do.par = FALSE,
                               cell.border.alpha = 0.1,
                               n.cores = 60) 
dev.off()

########## scVelo for latent time 
# This workflow assumes you already are familiar with R based scrna-seq analysis and biowulf
# Export Seurat object data, running the below code in R
# save metadata table:
eyelatentime <- eyecells
eyelatentime$barcode <- colnames(eyelatentime)
eyelatentime$UMAP_1 <- eyelatentime@reductions$umap@cell.embeddings[,1]
eyelatentime$UMAP_2 <- eyelatentime@reductions$umap@cell.embeddings[,2]
write.csv(eyelatentime@meta.data, file='eyelatentimemetadata.csv', quote=F, row.names=F)

# write expression counts matrix
library(Matrix)
counts_matrix <- GetAssayData(eyelatentime, assay='RNA', slot='counts')
writeMM(counts_matrix, file='eyelatentimecounts.mtx')    
# write dimesnionality reduction matrix, in this example case pca matrix
write.csv(eyelatentime@reductions$pca@cell.embeddings, file='eyelatentimepca.csv', quote=F, row.names=F)
# write gene names
write.table(data.frame('gene'=rownames(counts_matrix)),file='eyelatentimegene_names.csv',quote=F,row.names=F,col.names=F)
```

##### 19. RNA velocity analysis (latenttime)
```{python scVelo, message=FALSE, warning=FALSE}
# Below is the python code, run in python
# Request for an interactive session in biowulf
module load scvelo
python
import scanpy as sc
import anndata
from scipy import io
from scipy.sparse import coo_matrix, csr_matrix
import numpy as np
import os
import pandas as pd

# load sparse matrix:
X = io.mmread("eyelatentimecounts.mtx")
# create anndata object
adata = anndata.AnnData(
  X=X.transpose().tocsr())
# load cell metadata:
cell_meta = pd.read_csv("eyelatentimemetadata.csv")
# load gene names: 
with open("eyelatentimegene_names.csv", 'r') as f:
  gene_names = f.read().splitlines()
# set anndata observations and index obs by barcodes, var by gene names
adata.obs = cell_meta
adata.obs.index = adata.obs['barcode']
adata.var.index = gene_names
# load dimensional reduction:
pca = pd.read_csv("eyelatentimepca.csv")
pca.index = adata.obs.index
# set pca and umap
adata.obsm['X_pca'] = pca.to_numpy()
adata.obsm['X_umap'] = np.vstack((adata.obs['UMAP_1'].to_numpy(), adata.obs['UMAP_2'].to_numpy())).T
# plot a UMAP colored by sampleID to test:
sc.pl.umap(adata, color=['RNA_snn_res.0.2'], frameon=False, save=True)

# save dataset as anndata format
adata.write('my_dataeye.h5ad')

# reload dataset
adata = sc.read_h5ad('my_dataeye.h5ad')
# import scvelo and other modules
import scvelo as scv
import scanpy as sc
import numpy as np
import pandas as pd
import anndata as ad
# analysis settings
scv.settings.verbosity = 3
scv.settings.set_figure_params('scvelo', facecolor='white', dpi=100, frameon=False)
# load loom files for spliced/unspliced matrices for each sample:
ldata_s1h1 = scv.read('/data/CaspiRNAseqData/Results/velocyto/sample1_hto1/sample_alignments.loom', cache=True)
ldata_s1h2 = scv.read('/data/CaspiRNAseqData/Results/velocyto/sample1_hto2/sample_alignments.loom', cache=True)
ldata_s1h3 = scv.read('/data/CaspiRNAseqData/Results/velocyto/sample1_hto3/sample_alignments.loom', cache=True)
ldata_s1h4 = scv.read('/data/CaspiRNAseqData/Results/velocyto/sample1_hto4/sample_alignments.loom', cache=True)
ldata_s3h1 = scv.read('/data/CaspiRNAseqData/Results/velocyto/sample3_hto1/sample_alignments.loom', cache=True)
# preprocess barcodes to match seurat object
testlist=ldata_s1h1.obs.index.tolist()
testlist=[bc.split(':')[1] for bc in testlist]
testlist=[bc.split('x')[0] for bc in testlist]
barcodes = [sub + '-S1_HTO_1' for sub in testlist]
ldata_s1h1.obs.index = barcodes
testlist=ldata_s1h2.obs.index.tolist()
testlist=[bc.split(':')[1] for bc in testlist]
testlist=[bc.split('x')[0] for bc in testlist]
barcodes = [sub + '-S1_HTO_2' for sub in testlist]
ldata_s1h2.obs.index = barcodes
testlist=ldata_s1h3.obs.index.tolist()
testlist=[bc.split(':')[1] for bc in testlist]
testlist=[bc.split('x')[0] for bc in testlist]
barcodes = [sub + '-S1_HTO_3' for sub in testlist]
ldata_s1h3.obs.index = barcodes
testlist=ldata_s1h4.obs.index.tolist()
testlist=[bc.split(':')[1] for bc in testlist]
testlist=[bc.split('x')[0] for bc in testlist]
barcodes = [sub + '-S1_HTO_4' for sub in testlist]
ldata_s1h4.obs.index = barcodes
testlist=ldata_s3h1.obs.index.tolist()
testlist=[bc.split(':')[1] for bc in testlist]
testlist=[bc.split('x')[0] for bc in testlist]
barcodes = [sub + '-S3_HTO_1' for sub in testlist]
ldata_s3h1.obs.index = barcodes

# make variable names unique
ldata_s3h1.var_names_make_unique()
ldata_s1h1.var_names_make_unique()
ldata_s1h2.var_names_make_unique()
ldata_s1h3.var_names_make_unique()
ldata_s1h4.var_names_make_unique()
# concatenate the three loom
ldatan= ldata_s1h1.concatenate([ldata_s1h2,ldata_s1h3, ldata_s1h4, ldata_s3h1])
# cleanup names
scv.utils.clean_obs_names(adata)
scv.utils.clean_obs_names(ldatan)
# merge matrices into the original adata object
adata = scv.utils.merge(adata, ldatan)

# plot umap to check
sc.pl.umap(adata, color='RNA_snn_res.0.2', frameon=False, legend_loc='on data', title='', save='_0_20.pdf')
# pre-process
scv.pp.filter_and_normalize(adata)
scv.pp.moments(adata)

# compute velocity 
scv.tl.velocity(adata, mode='stochastic')
scv.tl.velocity_graph(adata)
# visualization
scv.pl.velocity_embedding_grid(adata, basis='umap', color='RNA_snn_res.0.2', save='embedding_grid.png', title='', scale=0.25,dpi = 300,
                               palette={'0':'#bfa708','2':'#48c785','3':'#33a7f5','4':'#d272ed'})
scv.pl.velocity_embedding_stream(adata, basis='umap', color='RNA_snn_res.0.2', save='embedding_stream0731.pdf', title='',dpi = 300,legend_loc='none',
                                 palette={'0':'#bfa708','2':'#48c785','3':'#33a7f5','4':'#d272ed'})
# latent time, takes time (overnight)
scv.tl.recover_dynamics(adata) 

#### 
scv.tl.latent_time(adata)
scv.pl.scatter(adata, color='latent_time', color_map='viridis', size=30, dpi = 300)
scv.pl.scatter(adata, color='latent_time', color_map='viridis', size=40, save='latenteye.pdf')

# save dataset as anndata format
adata.write('my_data_lantenteye.h5ad')

# Reference
# https://smorabit.github.io/tutorials/8_velocyto/
# https://www.10xgenomics.com/resources/analysis-guides/trajectory-analysis-using-10x-Genomics-single-cell-gene-expression-data
# https://scvelo.readthedocs.io/en/stable/
```

```{r save, message=FALSE, warning=FALSE}
# Save the modified R environment into a new RData filein the working directory
save.image(file=paste(outputpath,"SeuratObject.RData",sep=""))

sessionInfo()
```
