
---
# This RMD file loads the cellranger count data and performs single-cell gene expression analysis
title: "Single-Cell Gene Expression Analysis Workflow"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: no
---
  
```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
options(knitr.duplicate.label = "allow")
knitr::opts_chunk$set()
```


## 1. Introduction
This document takes the cellranger count data for each of the samples, does qc anlaysis, filtering, generates before and after qc plots and combines the data.
```{r packages, message=FALSE, warning=FALSE}
# Install hdf5 library if its not already installed
#sudo apt-get install libhdf5-dev or brew install hdf5
#install.packages("hdf5r")
# Load below libraries
library(Seurat)
library(ggplot2)
library(cowplot)
library(tidyverse)
```

## 2. Data setup
```{r datasetup, message=FALSE, warning=FALSE}
# Set the input, output paths
directorypath="/Users/nagarajanv/OneDrive - National Institutes of Health/Zixuan/Results/cellranger/"
outputpath="/Users/nagarajanv/OneDrive - National Institutes of Health/Zixuan/Results/"
sampleinfopath="/Users/nagarajanv/OneDrive - National Institutes of Health/Zixuan/SampleInformation.tab"
samples=c("S1_HTO_1","S1_HTO_2","S1_HTO_3","S1_HTO_4","S2_HTO_1","S2_HTO_2","S2_HTO_3","S2_HTO_4")
setwd("/Users/nagarajanv/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/Zixuan/Results/")
```

## 3. Quality Analysis and Filtering
This part of the workflow reads in each of the samples, analyses the quality, performs the quality filter and merges the data. Both merged unfiltered and merged filtered data are then used for plotting the quality
```{r qc, message=FALSE, warning=FALSE}
# Using the for loop to reads the samples list
# Create empty lists to hold the combined data set
allsamplestomerge=list()
allsamplestomergeprefilter=list()

# Using the for loop to reads the samples list
for(i in samples)
{
  # Remove any old variables
  rm(onesampledataFilt)
  rm(onesampledata)
  rm(onesample)
  # Print the current sample name
  print(i)
  # Read the data file
  onesample <- Read10X_h5(paste0(directorypath,paste0(i,"/sample_filtered_feature_bc_matrix.h5",sep=""),sep=""), use.names = T)

  # Print column names
  colnames(onesample$`Gene Expression`)
  # Add gene names to columnnames
  colnames(onesample$`Gene Expression`) <-     paste(sapply(strsplit(colnames(onesample$`Gene Expression`),split="-"),'[[',1L),i,sep="-")
  # Print column names
  colnames(onesample$`Gene Expression`)
  
  # Create a Seurat data object from the gex matrix, keeping all the features, and only cells with at least 300 genes
  onesampledata <- CreateSeuratObject(counts = onesample[["Gene Expression"]], min.cells = 0,min.features = 300,names.field = 2,names.delim = "\\-")
  
  # Check genes names
  head(rownames(onesampledata))
  # Check data rows and columns
  head(onesampledata)
  
  # Generate Percent mito for each of the cells
  # Lookup genes
  grep("^mt-",rownames(onesampledata@assays$RNA@counts),value = TRUE)
  grep("Rp[sl]",rownames(onesampledata@assays$RNA@counts),value = TRUE)
  grep("^Hb[^(p)]",rownames(onesampledata@assays$RNA@counts),value = TRUE)
  
  # Generate Percent mito for each of the cells
  onesampledata$percent.mito <- PercentageFeatureSet(onesampledata, pattern = "^mt-")
  summary(onesampledata$percent.mito)
  head(onesampledata)
  
  # Generate Percent ribo for each of the cells
  onesampledata$percent.ribo <- PercentageFeatureSet(onesampledata, pattern = "Rp[sl]")
  summary(onesampledata$percent.ribo)
  head(onesampledata)
  
  # Generate Percent hemoglobin for each of the cells
  onesampledata$percent.hb <- PercentageFeatureSet(onesampledata, pattern = "^Hb[^(p)]")
  summary(onesampledata$percent.hb)
  head(onesampledata)
  
  # Check the total number of cells in this sample
  table(onesampledata$orig.ident)
  # Filter data and keep only cells that has percent.mito less than or equal to 8
  onesampledataFilt <- subset(onesampledata, percent.mito <= 8)
  # Check the number of cells after mito filtering
  table(onesampledataFilt$orig.ident)
  # Filter for nCount RNA (total rna molecules sequenced)
  onesampledataFilt <- subset(onesampledataFilt, nCount_RNA >= 1000 & nCount_RNA <= 40000)
  # Check number of cells after ncount filtering
  table(onesampledataFilt$orig.ident)
  # Filter for nfeature (total number of genes expressed)
  onesampledataFilt <- subset(onesampledataFilt, nFeature_RNA >= 400)
  # Check number of cells after nfeature filtering
  table(onesampledataFilt$orig.ident)
  # Filter for percent ribo
  onesampledataFilt <- subset(onesampledataFilt, percent.ribo > 10 & percent.ribo < 45)
  # Counts after filtering
  table(onesampledataFilt$orig.ident)
  
  # Add filtered and unfiltered data to a list
  allsamplestomerge[[i]]=onesampledataFilt
  allsamplestomergeprefilter[[i]]=onesampledata

}
```

## 4. Merge samples
```{r merge, message=FALSE, warning=FALSE}
# Create one and the rest of the samples as a list, for filtered data
one=allsamplestomerge[[1]]
one
rest=allsamplestomerge[2:length(allsamplestomerge)]
rest

# Create one and the rest of the samples as a list, for prefiltered data
onepre=allsamplestomergeprefilter[[1]]
onepre
restpre=allsamplestomergeprefilter[2:length(allsamplestomergeprefilter)]
restpre

# Merge filtered seurat objects
allsamples <- merge(one, y = rest)
allsamples
head(allsamples)
table(Idents(allsamples))

# Merge prefiltered seurat objects
allsamplespre <- merge(onepre, y = restpre)
allsamplespre
head(allsamplespre)
table(Idents(allsamplespre))

```

## 5. Add metadata
```{r metadata, message=FALSE, warning=FALSE}
# Assign sample types metadata column
allsamples$type <- plyr::mapvalues(
  x = allsamples$orig.ident, 
  from = c("S1_HTO_1","S1_HTO_2","S1_HTO_3","S1_HTO_4","S2_HTO_1","S2_HTO_2","S2_HTO_3","S2_HTO_4"),
  to = c("S1","S1","S1","S1","S2","S2","S2","S2")
)
head(allsamples)

allsamplespre$type <- plyr::mapvalues(
  x = allsamplespre$orig.ident, 
  from = c("S1_HTO_1","S1_HTO_2","S1_HTO_3","S1_HTO_4","S2_HTO_1","S2_HTO_2","S2_HTO_3","S2_HTO_4"),
  to = c("S1","S1","S1","S1","S2","S2","S2","S2")
)
head(allsamplespre)

# View type counts
table(allsamples$type)
table(allsamplespre$type)
# View data
head(allsamples)


```

## 6. Generate before and after QC plots
```{r plots, message=FALSE, warning=FALSE}
# Generate before filtering quality plots
qcbefore=VlnPlot(allsamplespre,features = c("nFeature_RNA", "nCount_RNA","percent.mito","percent.ribo"),ncol = 2, pt.size = 0) +
  NoLegend()
qcbefore
# Save the plot in current working directory
ggsave(paste(outputpath,"before-qc-violineplot.pdf",sep=""), plot = qcbefore, width = 15, height = 20, units = "cm")

# Generate after filtering quality plots
qcafter=VlnPlot(allsamples,features = c("nFeature_RNA", "nCount_RNA","percent.mito","percent.ribo"),ncol = 2, pt.size = 0) +
  NoLegend()
qcafter
# Save the plot in current working directory
ggsave(paste(outputpath,"after-qc-violineplot.pdf",sep=""), plot = qcafter, width = 15, height = 20, units = "cm")
```

## 7. Normalization
```{r normalize, message=FALSE, warning=FALSE}
###################### Normalize Data
# Feature counts for each cell are divided by the total counts for that cell and multiplied by the scale.factor. This is then natural-log transformed
# Cell level normalization - accounts for sequencing depth
allsamples <- NormalizeData(
  object = allsamples,
  normalization.method = "LogNormalize",
  scale.factor = 10000)

##################### Find variable genes
# Returns top 2000 variable genes
# VST is uses LOESS method
# FindVariableFeatures needs normalization
allsamples <- FindVariableFeatures(
  object = allsamples,
  selection.method = "vst")
top10 <- head(VariableFeatures(allsamples), 10)
top10
# Plot variable features and label the top 10
vfp1 <- VariableFeaturePlot(allsamples)
vfp1 <- LabelPoints(plot = vfp1, points = top10, repel = TRUE)
vfp1
ggsave(paste(outputpath,"variablefeatureplot.png",sep=""), plot = vfp1, width = 20, height = 15, units = "cm")

```

## 8. Scaling & Batch Correction
```{r scale, message=FALSE, warning=FALSE}
##################### SCALE DATA
# PCA needs scaled data
# Zero centers and scales (mean/sd) gene/feature data in each cell (for across sample comparison), so extreme ranges in expression do not affect clustering (done for making cells with similar expression cluster together)
dim(allsamples)
allsamples <- ScaleData(
  object = allsamples,
  # Remove unwanted sources of variation (technical, batch etc.)
  vars.to.regress = c("percent.mito", "nFeature_RNA", "percent.ribo", "nCount_RNA"))

```
## 9. PCA
```{r PCA, message=FALSE, warning=FALSE}
##################### PCA
# Run PCA with 50 pcs
allsamples <- RunPCA(object = allsamples, npcs=50)
# Print genes in the top 5 pcs
print(allsamples[["pca"]], dims = 1:5, nfeatures = 5)
# Identify number of PCs that explains majority of variations
ElbowPlot(allsamples, ndims=50, reduction = "pca")
# Visualize the genes in the first PC
VizDimLoadings(allsamples, dims = 1, ncol = 1) + theme_minimal(base_size = 8)
# Visualize the genes in the second PC
VizDimLoadings(allsamples, dims = 2, ncol = 1) + theme_minimal(base_size = 8)
# Set the identity column to WT vs TG
Idents(allsamples) <- "orig.ident"
# Plot heatmap with cells=500 plotting cells with extreme cells on both ends of spectrum
#DimHeatmap(object = allsamples, dims = 1:6, cells = 500, balanced = TRUE)
DimHeatmap(object = allsamples, dims = 1:6, cells = 500, balanced = TRUE)
```

## 10. Clustering
```{r clustering, message=FALSE, warning=FALSE}
##################### Clustering
# use the first 20 pc's
use.pcs = 1:20
# FindNeighbors with 20 pcs
allsamples <- FindNeighbors(allsamples, reduction="pca", dims = use.pcs)
# FindClusters with resolution starting at 0.25 and ending at 4, with 0.5 increment
allsamples <- FindClusters(
  object = allsamples,
  resolution = seq(0.20,0.65,0.15),
  verbose = FALSE
)
head(allsamples)
# Count number of clusters at each resolution
sapply(grep("res",colnames(allsamples@meta.data),value = TRUE),
       function(x) length(unique(allsamples@meta.data[,x])))
```

## 11. tSNE
```{r tsne, message=FALSE, warning=FALSE}
####################### TSNE
# Run tsne with 20 pcs
allsamples <- RunTSNE(
  object = allsamples,
  reduction.use = "pca",
  dims = use.pcs,
  do.fast = TRUE)

# Plot all clusters in all resolutions
DimPlot(object = allsamples, group.by=grep("res",colnames(allsamples@meta.data),value = TRUE)[1:4], ncol=2 , pt.size=3.0, reduction = "tsne", label = T)
# change default identity
Idents(allsamples) <- "RNA_snn_res.0.35"
# list cell number in each cluster for sample types
table(Idents(allsamples),allsamples$type)
# Visualize cluster at 0.35 resolution
DimPlot(object = allsamples, pt.size=0.5, reduction = "tsne", label = T)
# Color by S1 vs S2
DimPlot(object = allsamples, pt.size=0.5, reduction = "tsne", group.by = "type" )
```

## 12. UMAP
```{r umap, message=FALSE, warning=FALSE}
########################## UMAP
# Run umap with 20 pcs
set.seed(20)
allsamples <- RunUMAP(
  object = allsamples,
  reduction.use = "pca",
  dims = use.pcs)
# Plot umap
DimPlot(object = allsamples, pt.size=0.5, reduction = "umap", label = T)
# Color by S1 vs S2
DimPlot(object = allsamples, pt.size=0.5, reduction = "umap", group.by = "type")

```

## 13. Differential expression (S1 vs S2)
```{r differentials1vss2, message=FALSE, warning=FALSE}
########################### Marker genes between groups
# Current identity levels
levels(allsamples)
head(Idents(allsamples))
# Change identity to S1 vs S2
Idents(allsamples) <- "type"
head(Idents(allsamples))
levels(allsamples)
########################### tables
# Proportion and counts of cells in clusters and groups
prop.table(table(Idents(allsamples)))
table(Idents(allsamples))
# Find differentially expressed genes for S1 vs S2
markerss1s2 = FindMarkers(allsamples, ident.1="S1", ident.2="S2")
head(markerss1s2)
# Set marker rownames
markerss1s2$gene=rownames(markerss1s2)
# Save the list of differentially expressed genes as a tab delimited text file
write.table(markerss1s2, file=paste(outputpath,"markerss1s2.txt",sep=""), quote=F, row.names = FALSE, sep = "\t")

VlnPlot(allsamples,features=c("Il17a","Il1r1"), pt.size = 0)
FeaturePlot(allsamples, features="Il17a")
table(allsamples$type)

```

## 14. Differential expression between clusters
```{r differentialbetweenclusters, message=FALSE, warning=FALSE}
########################### tables
# Proportion and counts of cells in clusters and groups
# Change identity to 0.35 cluster resolution
Idents(allsamples) <- "RNA_snn_res.0.35"
prop.table(table(Idents(allsamples)))
table(Idents(allsamples), allsamples$type)

########################### Marker genes between clusters
# Current identity levels
levels(allsamples)
head(Idents(allsamples))

# Find differentially expressed genes for cluster 0 vs 1
markerscluster = FindMarkers(allsamples, ident.1=0, ident.2=1)
head(markerscluster)
write.table(markerscluster, file=paste(outputpath,"markerscluster.txt",sep=""), quote=F, row.names = FALSE, sep = "\t")

VlnPlot(allsamples,features=c("Il17a","Il18r1"), pt.size = 0)
FeaturePlot(allsamples, features="Il18r1")
table(allsamples$orig.ident)

```
## 15. Differential expression between every clusters
```{r differentialbetweeneveryclusters, message=FALSE, warning=FALSE}
###########################
markersall=FindAllMarkers(allsamples)
head(markersall)
#View(markersall)
write.table(markersall, file=paste(outputpath,"markersall.txt",sep=""), quote=F, row.names = FALSE, sep = "\t")

VlnPlot(allsamples,features=c("Il17a","Nrgn"), pt.size = 0)
FeaturePlot(allsamples, features="Nrgn")

# Extract the top marker for each cluster
markersall %>%
  group_by(cluster) %>%
    slice(1) %>%
      pull(gene) -> best.gene.per.cluster

best.gene.per.cluster

# Extract the top 10 marker for each cluster
head(markersall)
markersall %>%
  group_by(cluster) %>%
  slice(1:10) %>%
  select(p_val, avg_log2FC, gene) -> best.gene.per.cluster.table

best.gene.per.cluster.table
write.table(best.gene.per.cluster.table, file=paste(outputpath,"markerstop10all.txt",sep=""), quote=F, row.names = FALSE, sep = "\t")

# Store plot as ggplot list
vlnplot1<-VlnPlot(allsamples,features=best.gene.per.cluster, pt.size = 0, combine = FALSE)
# vlnplot1<-VlnPlot(allsamples,features=best.gene.per.cluster, pt.size = 0, same.y.lims = TRUE, combine = FALSE)

# Set properties for each plot in the list
p1 <- list()
for (i in seq_along(vlnplot1)){
    #Change x and y tick label font size.
    p1[[i]] = vlnplot1[[i]] +
      theme(legend.position = "none",
            axis.text.x = element_text(size = 8),
            axis.text.y = element_text(size = 8),
            axis.title.y = element_text(size = 6),
            axis.title.x = element_blank(),
            title = element_text(size=8))
}
# Plot all the plots in the list
plot_grid(plotlist = p1, ncol = 3)

```

## 16. Differential expression groups within a specific cluster

```{r differentialgroupsinacluster, message=FALSE, warning=FALSE}
########################### tables
# Proportion and counts of cells in clusters and groups
# Change identity to 0.35 cluster resolution
Idents(allsamples) <- "RNA_snn_res.0.35"
prop.table(table(Idents(allsamples)))
table(Idents(allsamples))

# Current identity levels
levels(allsamples)
head(Idents(allsamples))

# Take all cells in cluster 0, and find markers that separate 'S1' and 'S2' that are in the 'type' column. subset.ident would the default identity of clusters
markersingroup <- FindMarkers(allsamples, ident.1 = "S1", ident.2 = "S2", group.by = 'type', subset.ident = 0)

head(markersingroup)
# Save markers
write.table(markersingroup, file=paste(outputpath,"markersingroup.txt",sep=""), quote=F, row.names = FALSE, sep = "\t")
            
VlnPlot(allsamples,features=c("Il17a","Cyp2j6"), group.by = "type", pt.size = 0)
FeaturePlot(allsamples, features=c("Foxp3"), split.by = "type")
table(allsamples$type)

```

## 17. Differential expression between mouse ear and S2

```{r differentialearvss2, message=FALSE, warning=FALSE}
########################### tables
# Proportion and counts of cells in ear and s2
# Change identity to 0.35 cluster resolution
Idents(allsamples) <- "RNA_snn_res.0.35"
prop.table(table(Idents(allsamples)))
table(Idents(allsamples))

# Current identity levels
levels(allsamples)
head(Idents(allsamples))

# Take all cells in cluster 4, and find markers that separate it from all cells in S2 (1,2,3,6,7,8). subset.ident would the default identity of clusters
markersinearvss2 <- FindMarkers(allsamples, ident.1 = 4, ident.2 = c(1,2,3,6,7,8))

head(markersinearvss2)
# Save markers
write.table(markersinearvss2, file=paste(outputpath,"markersinearvss2.txt",sep=""), quote=F, row.names = FALSE, sep = "\t")
            
VlnPlot(allsamples,features=c("Il1r1","Ikzf2"), group.by = "type", pt.size = 0)
FeaturePlot(allsamples, features=c("Il1r1"), split.by = "type")
table(allsamples$type)

```

```{r reference}
#https://ucdavis-bioinformatics-training.github.io/2021-March-Single-Cell-RNA-Seq-Analysis/data_analysis/scRNA_Workshop-PART4_fixed
```



```{r save, message=FALSE, warning=FALSE}
# Save QC session data in the working directory
save.image(file=paste(outputpath,"SeuratObjectAfterMarkerDiscovery.RData",sep=""))

sessionInfo()
```

